---
title: "Statistical analysis plan for calibration of the Trident trial to a population-based cohort in Scotland."
author: "Dr Alice Hosking, Prof David McAllister, ...."
date: "`r Sys.Date()`"
output: 
  html_document:
            toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(flexsurv)
library(survival)
library(SurvRegCensCov)
library(muhaz)
library(kableExtra)
library(ggsurvfit)
library(survminer)
library(rmeta)

set.seed(42)

knitr::opts_chunk$set(message = F, warning=F, echo=F)
```

## Introduction

### Purpose and scope of this document

This statistical analysis plan details the analyses planned for calibration of the TRIDENT trial to a population-based inception cohort, including a worked-through example with mock values of hazard ratios.

### Study overview

Intracerebral haemorrhage causes around a third of strokes worldwide but half the death and disability after stroke.1 Those with ICH have an estimated 2% annual risk of recurrent ICH and 2% annual risk of ischaemic stroke.2 A single-pill combination of three blood pressure lowering medicines at half standard dose (telmisartan 20 mg, amlodipine 2.5 mg, and indapamide 1.25 mg) in addition to standard care could reduce the risk of  recurrent stroke after ICH. TRIDENT was an international, double-blind, placebo-controlled, randomized trial in adults with ICH and mild-moderate hypertension (systolic BP: 130–160 mmHg. The primary outcome was any stroke; secondary outcomes include recurrent ICH, cardiovascular events, and safety.3 

However, TRIDENT, as with any other trial, might be subject to selection bias, leading to low representativity of the general population (i.e., lack of external validity).4–6 In particular, women, vulnerable older people, and those with frailty and multimorbidity, are likely underrepresented in this trial.7–9 Furthermore, the trial was conducted in countries with different demography to the UK.

In recent years statistical trial calibration (or transportation), has been developed to deal with this issue.10 Calibration refers to the methods used to extend the results of a clinical trial to a different population than the one originally studied. Rather than using real-world cohort studies to assess the same clinical question, calibration keeps the randomisation intact and therefore does not introduce confounding bias. This process provides a quantitative estimate of the external validity of trial findings.


### Objectives

The aim of this study is to calibrate the effect of a low-dose triple pill antihypertensive medication in people with ICH from the TRIDENT trial to a population-based cohort in Lothian, Scotland.


## Study design, population and outcomes

### Study design

This study will use stratified data from the TRIDENT trial (the trial population) and data from the LATCH study (the target population) a prospective, population-based inception cohort study of people with ICH in NHS Lothian. 

The TRIDENT trial result showed that there appeared to be no difference in the primary outcome (recurrent stroke) until approximately 18 months, from which point the cumulative incidence function for the primary outcome appears to diverge between control and treatment arms. We will therefore calibrate at two time points, 18 months ($t_1$) and end of follow up (approx 6 years, $t_1+t_2=t_3$).

In summary, the method is:

* Inspect the distribution of ages in TRIDENT and LATCH to choose sensible stratification ages.
* Split TRIDENT into strata of age and sex (eg <65 years female, <65 years male, over 65 years female,  over 65 years male)
* Obtain hazard ratios for TRIDENT for each stratum at 18 months and 4.5 years. 
* Split LATCH into the same age- and sex-based strata. For each strata, fit a survival model to obtain rates and confidence intervals of the primary outcome at t1 and t3.
* Use these rates, and the hazard ratios from TRIDENT, to model absolute rates in untreated and treated in each stratum.
* The relative risk for each stratum is then (risk untreated/risk treated). We combind relative risks and use simulation to obtain confidence intervals.

### Populations

#### Trial population

The trial population is the 1500 people included in the TRIDENT trial, an international multicentre double-blind placebo-controlled trial. The average time to randomisation was 54 days.

TRIDENT Inclusion Criteria
* Adults (≥18 years) with a history of primary ICH that was confirmed by imaging.
* Clinically stable, as judged by investigator.
* An average of two resting systolic blood pressure (SBP) levels measured 5 minutes apart in the range 130-160 mmHg recorded in a seated position. Patients with higher SBP could still be included if it was considered by the site investigator that their management would be consistent with local standards of clinical practice.
* Geographical proximity to the recruiting hospital and/or follow-up medical clinic site to allow ready access for in-person assessments during follow-up.
* No clear contraindication or indication to any of the individual components of the Investigational Medical Product (IMP).
* Provision of written informed consent
TRIDENT Exclusion Criteria
* Taking an angiotensin converting enzyme inhibitor (ACE-I) that could not be switched to any of the following alternatives: an angiotensin receptor blocker (ARB) (e.g. telmisartan 20 or 40 mg), a calcium channel blocker (CCB) (e.g. amlodipine 2.5 or 5 mg), or a thiazide (TZ)-like diuretic (e.g. indapamide 1.25 mg); an equivalent class (ARB, CCB or TZ); or a beta-blocker (BB).
* Contraindication to any component of the IMP in the context of currently prescribed BP lowering treatment.
* Unable to complete the study procedures and/or follow-up.

#### Target population

The Lothian Audit of Intracerebral Haemorrhage (LATCH) ran between 1st June 2010 and 31st May 2013, and 1st November 2018 and 31st October 2021. Ethical approval for LATCH was granted by the Caldicott guardian of NHS Lothian. The LATCH audit has been described previously.11 In brief, LATCH ascertained cases of first-ever or recurrent ICH from the Lothian Health board region of Scotland in residents aged ≥16 years. Prospective sources included a trust-wide network of physicians, neurologists, neurosurgeons, radiologists, pathologists, specialist nurses and audit personnel; and daily neuroradiology meetings and review of all computed tomographic (CT) head scans. We conducted retrospective searches of electronic patient notes, International Classification of Diseases 10th Revision (ICD-10) coded hospital discharge records, the Scottish Stroke Care Audit,12 and records at the Office of the Procurator Fiscal.  

The LATCH cohort includes 1077 people with ICH of any cause; 917 had ICH presumed due to microvascular cause. The LATCH study does not include information on clinical stability so we make assumptions to mirror the TRIDENT inclusion criteria. We aim to identify those in LATCH who might be prescribed the pill, if it were recommended in clinical treatment guidelines.

*	In the LATCH cohort, a third of people die within 7 days of ICH; beyond 7 days, the death rate reduces, and we pragmatically assume those alive at 7 days after ICH are clinically stable; there are 602 people alive at 7 days. 
*	If the triple pill used in TRIDENT were used in clinical practice, there would likely be no upper restriction on pre-treatment blood pressure; therefore we do not limit by blood pressure. 
*	The pill would only be used on those able to get out of bed; therefore, we will limit inclusion to those with pre-morbid modified Rankin Scale (mRS) of 4 or less. All 602 who survived 7 days had a pre-morbid mRS of 4 or less.

#### Overlap between trial and target populations.

The trial and target populations are partially overlapping, as the Royal Infirmary of Edinburgh was recruited 17 people with ICH for TRIDENT. This does not affect the validity of the process of calibration.

## Analysis example

We now show the method for calibration. We use LATCH data. We create four strata for this example, and mock hazard ratios for TRIDENT for these strata.

### Data tidying

First we filter LATCH to only those who survived 7 days, format dates correctly, set censoring dates, and split into strata (under 65 and 65 and over, male and female). For this study, we censor LATCH follow up at the maximum TRIDENT follow-up time of six years. (Code not shown.)

```{r read in data, echo=FALSE, message=FALSE}
latch_for_tridentx<- read_csv("/Users/alicehosking/Documents/REA3 clinical fellow job/Treatment effect in trials/LATCH cohort survived 7 days.csv")
setDT(latch_for_tridentx)
latch_for_trident<- latch_for_tridentx[,c("survival_time","censor_status_death","i_date","Gender_t", "date_of_death.y","cause_of_death","first_mi","first_tia","first_stroke","first_dvt","first_pe", "first_rec_ich","pre_rankin", "age_index")] 
#some dates are stored as excel formatted dates. 
#for this analysis, we care about date of first stroke (which is ischaemic stroke) and date of recurrent ICH
latch_for_trident[, first_rec_ich:= as.Date(first_rec_ich, origin = "1899-12-30")][, first_stroke:= as.Date(first_stroke, origin = "1899-12-30")]

#time to first event
latch_for_trident[, time_to_ich:= difftime(first_rec_ich,i_date,  units="days")][,time_to_isch_str:=difftime(first_stroke ,i_date,  units="days")][, time_to_all_stroke:= pmin(time_to_ich, time_to_isch_str, na.rm=T)][,date_of_death_trid:= as.Date(date_of_death.y, origin = "1899-12-30")] [, time_to_death:= difftime(date_of_death_trid,i_date,  units="days")]

# if no first event, time to end of death/ end study
latch_for_trident[, surv_time_all_stroke:=fifelse(is.na(time_to_all_stroke)==T&is.na(time_to_death)==T, difftime("2022-12-31", i_date, units="days"), fifelse(is.na(time_to_all_stroke)==T&is.na(time_to_death)==F, time_to_death, time_to_all_stroke))]

#censor status and survival time at 1.5 years, 4.5 years and all time
latch_for_trident[, censor_status_all_str_all_time:= fifelse(is.na(time_to_all_stroke)==T,0,1)][,censor_all_str_t1:= fifelse(time_to_all_stroke<548,1,0)][ ,censor_all_str_t1:=fifelse(is.na(time_to_all_stroke)==T,0,censor_all_str_t1)][, survival_stroke_t1:= fifelse(censor_all_str_t1==1, time_to_all_stroke, pmin(surv_time_all_stroke, time_to_death, na.rm = T))][, survival_stroke_t1:= pmin(548,survival_stroke_t1)][,censor_all_str_t3:= fifelse(time_to_all_stroke<6*365.25,1,0)][ ,censor_all_str_t3:=fifelse(is.na(time_to_all_stroke)==T,0,censor_all_str_t3)][, survival_stroke_t3:= fifelse(censor_all_str_t3==1, time_to_all_stroke, pmin(surv_time_all_stroke, time_to_death, na.rm = T))][,survival_stroke_t3:=pmin(round(6*365.25,0),survival_stroke_t3)]

strata<- list(
latch_under_65_m<- latch_for_trident[age_index<65&Gender_t=="Male"],
latch_under_65_f<- latch_for_trident[age_index<65&Gender_t=="Female"],
latch_over_65_m<- latch_for_trident[age_index>64&Gender_t=="Male"],
latch_over_65_f<- latch_for_trident[age_index>64& Gender_t=="Female"]
  
)

strata_n<- c()
strata_n_events_t1<- c()
strata_n_events_t3<- c()
strata_fu_t1 <- c()
strata_fu_t3<- c()

for(i in 1:length(strata)){
  strata_n[i] <-length(strata[[i]]$i_date) 
  strata_n_events_t1[i]<- sum(strata[[i]]$censor_all_str_t1)
  strata_n_events_t3[i]<- sum(strata[[i]]$censor_all_str_t3)
  strata_fu_t1[i] <- sum(strata[[i]]$survival_stroke_t1)
  strata_fu_t3[i]<- sum(strata[[i]]$survival_stroke_t3)

}

```


### Parametric survival model for LATCH {.tabset}

We model the event rate (stroke recurrence) in LATCH for each stratum. We use a parametric model. 

First inspect the Kaplan Meier curve. This looks approximately exponential. 

We model the event rate with a parametric survival model to select the most appropriate distribution, using the flexsurv package in R. There are various choices for the model distribution (exponential, gamma, Weibull etc). We compare the hazard and survival plots for each distribution to a non-parametric model (the kernel, in black), and the AIC for each distribution.

The exponential model allows much simpler calculations other models. The AICs are all similar (gamma is the lowest). Although other models have marginally better fits, the exponential model fit is acceptable, and for analytical simplicity, we choose an exponential model.


#### Kaplan Meier curve


```{r KM for whole population, warnings=FALSE, message=FALSE}

s1<- survfit(Surv(survival_stroke_t3, censor_all_str_t3) ~ 1, data = latch_for_trident)
km_plot<- survminer::ggsurvplot(s1, ylim=c(0.6, 1), xlim=c(0, 6*365.25), risk.table =FALSE,surv.plot.height = 2,
           ggtheme = theme(aspect.ratio = 1),legend="none")
km_plot
```



```{r choose parametric model for event rate}

# help here: https://www.r-bloggers.com/2019/06/parametric-survival-modeling/
#from help page - choosing best model based on shape of distribution

# cran vignette from flexsurv package also useful
# https://cran.r-project.org/web/packages/flexsurv/vignettes/standsurv.html

#non-parametric model for comparison
kernel_haz_est <- muhaz(as.difftime(latch_for_trident$surv_time_all_stroke, units="days"), latch_for_trident$censor_status_all_str_all_time)
kernel_haz <- data.table(time = kernel_haz_est$est.grid, est = kernel_haz_est$haz.est,method = "Kernel density")

#vectors of all the options
dists <- c("exp", "weibull",  "gamma", "lognormal", "llogis", "gengamma","gompertz")
dists_long <- c("Exponential", "Weibull (AFT)", "Gamma", "Lognormal", "Log-logistic", "Generalized gamma","Gompertz")

#parametric model - generate hazards and survival for different distributions, for whole dataset
#initialise variables
parametric_haz <- vector(mode = "list", length = length(dists))
parametric_preds <- vector(mode = "list", length = length(dists))
fit_test<- list()

#loop through each distribution
for (i in 1:(length(dists)-1)){
  fit <- flexsurvreg(Surv(survival_stroke_t3, censor_all_str_t3) ~ 1, data = latch_for_trident, dist = dists[i]) 
  fit_test[[i]]<- fit
  #hazards for plot 1
  parametric_haz[[i]] <- summary(fit, type = "hazard", ci = FALSE, tidy = TRUE)
  parametric_haz[[i]]$method <- dists_long[i]
  #survival for plot 2
  parametric_preds[[i]] <- summary(fit, type = "survival", tidy=T)
  parametric_preds[[i]]$method <- dists_long[i]
  
  }

#add in gompertz separately as failed to initialize
fit_gompertz <- flexsurvreg(Surv(survival_stroke_t3, censor_all_str_t3) ~ 1, data = latch_for_trident, dist = dists[7], inits=c(-1, 1/mean(as.numeric(latch_for_trident$surv_time_all_stroke)))) 
parametric_haz[[length(dists)]] <- summary(fit_gompertz, type = "hazard", 
                                     ci = FALSE, tidy = TRUE)
parametric_haz[[length(dists)]]$method<- dists_long[length(dists)]
parametric_preds[[length(dists)]] <- summary(fit_gompertz, type = "survival", 
                                     ci = FALSE, tidy = TRUE)
parametric_preds[[length(dists)]]$method<- dists_long[length(dists)]

#join all together
parametric_haz <- rbindlist(parametric_haz)
haz <- rbind(kernel_haz, parametric_haz)
haz[, method := factor(method,
                       levels = c("Kernel density",
                                  dists_long))]
n_dists <- length(dists) 
```
#### Plot of survivals of various distributions

```{r survival plots}
#alternative plot
parametric_preds_df<- rbindlist(parametric_preds, fill=TRUE)
ggplot() + geom_line(aes(x=time, y=est, col = method, linetype = method ), data=parametric_preds_df) + 
  geom_step(aes(x=time, y=surv), data=km_plot$data.survplot)
```

#### Plot of hazards of various distributions
```{r hazard plot}

ggplot(haz, aes(x = time, y = est, col = method, linetype = method)) +
  geom_line() +
  xlab("Days") + ylab("Hazard") + 
  scale_colour_manual(name = "", 
                      values = c("black", rainbow(n_dists))) +
  scale_linetype_manual(name = "",
                        values = c(1, rep_len(2:6, n_dists)))+
  xlim(0, 365.25*6)

```



#### Compare AIC

```{r aic}

parametric_haz <- vector(mode = "list", length = length(dists))
fit_test<- list()
fit_aic<- c()
for (i in 1:(length(dists)-1)){
  fit <- flexsurvreg(Surv(survival_stroke_t3, censor_all_str_t3) ~ 1, data = latch_for_trident, dist = dists[i]) 
  fit_test[[i]]<- fit
  fit_aic<- c(fit_aic, fit$AIC)
}

#add in gompertz separately
fit_gompertz <- flexsurvreg(Surv(survival_stroke_t3, censor_all_str_t3) ~ 1, data = latch_for_trident, dist = dists[7], inits=c(-1, 1/mean(as.numeric(latch_for_trident$time_to_all_stroke)))) 
fit_aic<-  c(fit_aic, fit_gompertz$AIC)
#join all together
parametric_aic<- as.data.frame(cbind(dists_long, round(fit_aic,2)))
kable(parametric_aic)
```


### Simple event rates {.tabset}

Assuming exponential event rate across t1 and t2, we can calculate the event rate, the standard error, and simulate a normally distributed sample. Throughout, we simulate with 10,000 samples. We plot the sample and show the event rates for each stratum in a table.

#### Plot of sample

```{r simple event rates}
## Set number of samples with nrep for use throughout.
nrep<-10000

#event rate at t1 and t2
rate_est <- strata_n_events_t3/ strata_fu_t3*365.35*100
log_rate_est <- log(strata_n_events_t3/ strata_fu_t3)
# Calculate SE of the log-rate for untreated
log_rate_se <- 1 / sqrt(strata_n_events_t3)
## obtain and plot samples from rate in untreated
untreated_rates<- as.data.frame(cbind(strata_n, rate_est,log_rate_est, log_rate_se))
setDT(untreated_rates)
log_lambda_u<- untreated_rates[,rnorm(nrep, mean=log_rate_est, sd=log_rate_se), by=.I]

#rename I column to the strata names
strata <-c("Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")
log_lambda_u$I <- factor(log_lambda_u$I, levels=c(1,2,3,4), labels=strata)

#plot to confirm normality
ggplot(data=log_lambda_u)+
  geom_histogram(aes(x=V1), fill="#FCCDE5")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated log rates for strata 1-4")+ 
  xlab("")
```

#### Log of event rates for each stratum

```{r table of rates}
#print table of rates
untreated_rates<- as.data.frame(untreated_rates)
rownames(untreated_rates)<- strata
colnames(untreated_rates) <- c( "N in stratum", "Events per 100 person years", "log(rate/day)", "log(standard error)")
(round(untreated_rates,2))%>%
  kable(row.names = T)%>%
  kable_styling()

```


### Hazard ratios for each stratum {.tabset}

For this mock-up we estimate hazard ratios for each time period (based roughly on published subgroup analysis). For the real analysis, the strata may be different; this is for demonstration only. We assume HR are independent; and make a guess at the HR for young men. We then split into two time periods: up to 18 months, where the HR are approximately 1; and beyond 18 months. We simulate 10,000 samples assuming a log normal distribution, for t1 and t2.


```{r mock hazard ratios - whole trial}
# overall HR. These come from the subgroup analysis of the trial
#create vector for each with HR, LCI, UCI
male_hr<- c(0.71,0.42,1.15)
female_hr<- c(0.44,0.21,0.94)
f_m_interaction <- female_hr/male_hr

under_60_hr<- c(0.76,0.44,1.33)
over_60_hr<- c(0.61,0.41,0.92)
old_young_interaction<- over_60_hr/under_60_hr

#assume independence; can multiply to get HR for subgroups
#need to make assumption on HR for young men

hrs_t2<- as.data.frame(rbind(
hr_young_men<- c(0.58, 0.46, 1.35),
hr_young_women <- hr_young_men*f_m_interaction,
hr_old_men<- hr_young_men*old_young_interaction,
hr_old_women <- hr_young_men*f_m_interaction*old_young_interaction
))
# we need to split into two time strata; the first 18 months and beyond 18 months. For the first 18 months, the HR between treatment and control overall is approximately 1; beyond this the HR widens. for the first 18 months, recentre HR and CI to 1ish. For beyond 18 months, use existing HR
hrs_t1<- as.data.frame(rbind(
hr_young_men_t1<- hr_young_men+0.42,
hr_young_women_t1 <- hr_young_women+0.72,
hr_old_men_t1 <- hr_old_men+0.68,
hr_old_women_t1 <- hr_old_women + 0.82
))

colnames(hrs_t1) <- c("HR", "l_ci", "u_ci")
rownames(hrs_t1) <- c(" Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")
colnames(hrs_t2) <- c("HR", "l_ci", "u_ci")
rownames(hrs_t2) <- c(" Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")
```

#### Hazard ratio for period t1
```{r hr for t1}
kable(round(hrs_t1,2))%>% kable_styling()
```

#### Hazard ratio for period t2

```{r hr for t2}
kable(round(hrs_t2,2))%>% kable_styling()
```

#### Histogram of simulated log hazard ratios for strata 1-4, T1

```{r simulate n hrs, warnings=F}
#obtain log of hrs and CI at t1 and t2. Assume log normally distributed, and find log standard error
setDT(hrs_t1)
setDT(hrs_t2)
hrs_t1[, log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]
hrs_t2[, log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]

#simulate log hr for nrep samples.
#create simulated hr for all strata
log_hrs_t1_per_strata <- hrs_t1[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]

#rename I column to the strata names
log_hrs_t1_per_strata$I <- factor(log_hrs_t1_per_strata$I, levels=c(1,2,3,4), labels=strata)


#plot to confirm normality
ggplot(data=log_hrs_t1_per_strata)+
  geom_histogram(aes(x=V1), fill="#FCCDE5")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated log hazard ratios for strata 1-4, T1")+ 
  xlab("Log of hazard ratio")
```

#### Histogram of simulated hazard ratios for strata 1-4, T1 

```{r hist hr t1 exponentiated}
ggplot(data=log_hrs_t1_per_strata)+
  geom_histogram(aes(x=exp(V1)), fill="#FCCDE5")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated hazard ratios for strata 1-4, T1")+ 
  xlab("Hazard ratio")
```

#### Histogram of simulated log hazard ratios for strata 1-4, T2

```{r t2 hr hists}
#repeat for t2 hazard ratios
#create simulated hr for all strata
log_hrs_t2_per_strata <- hrs_t2[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]

#rename I column to the strata names
log_hrs_t2_per_strata$I <- factor(log_hrs_t2_per_strata$I, levels=c(1,2,3,4), labels=strata)

#plot to confirm normality
ggplot(data=log_hrs_t2_per_strata)+
  geom_histogram(aes(x=V1), fill="#80B1D3")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated log hazard ratios for strata 1-4, T2")+ 
  xlab("Log of hazard ratio")
```

#### Histogram of simulated hazard ratios for strata 1-4, T2 

```{r t2 hr exponentiated}
ggplot(data=log_hrs_t2_per_strata)+
  geom_histogram(aes(x=exp(V1)), fill="#80B1D3")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated hazard ratios for strata 1-4, T2")+ 
  xlab("Hazard ratio")
```

### Calibrated rates and CIs for each stratum {.tabset}

For each age/sex stratum and for each time period, we combine the rate in the untreated and the hazard ratio to obtain a risk ratio.

We first define the following:

$$\lambda_u=\text{rate in untreated}$$
$$t_1=\text{length in days of first time period}$$
$$t_2=\text{length in days of second time period}$$

Multiply untreated rates by hazard ratios ($HR$) to get treated event rates:
$$\lambda_{t1} = \lambda_u \times HR_1$$
$$\lambda_{t2} = \lambda_u \times HR_2$$

We then transform these to risks for each stratum:
$$R_{untreated} = 1 - e^{-\lambda_u(t_1+t_2)}$$ 

and 

$$R_{treated} = 1 - e^{-(\lambda_{t1} t_1 + \lambda_{t2} t_2)}$$

The risk ratio at the end of the trial is then:

$$ RR= \frac{R_{treated}}{R_{untreated}}$$

We have already simulated 10,000 samples for each hazard ratio. We simulate 10,000 samples for event rates, which gives 10,000 risk ratios.

```{r risks treated and untreated 2}
#set t1 and t2 in days
t1<-548
t2<- 4.5*365.25

#lambda from parametric model above
#lambda_u <- exp(z_lambda) 
log_lambda_u_sim<- log_lambda_u$V1

#for rate in treated, log lambda_t = log lambda_u + log hr
log_lambda_t1_sim <- log_lambda_u_sim+log_hrs_t1_per_strata$V1
log_lambda_t2_sim <- log_lambda_u_sim+log_hrs_t2_per_strata$V1

#exponentiate to get rates
lambda_u_sim <- exp(log_lambda_u_sim)
lambda_t1_sim<- exp(log_lambda_t1_sim)
lambda_t2_sim<- exp(log_lambda_t2_sim)

#transform to risks
risk_u1 <- 1-exp(-lambda_u_sim*t1)
risk_t1 <- 1-exp(-lambda_t1_sim*t1)
risk_u2 <- 1-exp(-lambda_u_sim*(t1+t2))
risk_t2 <- 1-exp(-(lambda_t1_sim*t1+lambda_t2_sim*t2))
risk_ratio_t1 <- risk_t1/risk_u1
risk_ratio_t2 <- risk_t2/risk_u2

risk_ratio_for_plot_t1 <- as.data.frame(cbind(c(rep(1,nrep), rep(2, nrep), rep(3, nrep), rep(4, nrep)), risk_ratio_t1))
risk_ratio_for_plot_t2 <- as.data.frame(cbind(c(rep(1,nrep), rep(2, nrep), rep(3, nrep), rep(4, nrep)), risk_ratio_t2))

rr_means_t1<- risk_ratio_for_plot_t1%>%
  group_by(V1)%>%
  summarize(means_t1=round(mean(risk_ratio_t1),3))

rr_means_t2<- risk_ratio_for_plot_t2%>%
  group_by(V1)%>%
  summarize(means_t2=round(mean(risk_ratio_t2),3))

```

#### Simulated risk ratios for strata 1-4, T1 

```{r risk plot t1}
ggplot(data=risk_ratio_for_plot_t1)+
  geom_histogram(aes(x=risk_ratio_t1), fill="#FCCDE5")+
  geom_text(data=rr_means_t1, aes(x=1, y=150, label= paste("Mean=",means_t1)))+
  theme_bw()+
  facet_wrap(vars(V1))+
  labs(title="Simulated risk ratios for strata 1-4, T1")+ 
  xlab("Risk ratio")
```

#### Simulated risk ratios for strata 1-4, T1+T2

```{r risk plot t2}
ggplot(data=risk_ratio_for_plot_t2)+
  geom_histogram(aes(x=risk_ratio_t2), fill="#80B1D3")+
  geom_text(data=rr_means_t2, aes(x=1, y=150, label= paste("Mean=",means_t2)))+
  theme_bw()+
  facet_wrap(vars(V1))+
  labs(title="Simulated risk ratios for strata 1-4, T1+T2")+ 
  xlab("Risk ratio")

```

### Pooling risk ratios {.tabset}

The final step is to pool the risk ratio for each stratum. We take a weighted mean, where the risk ratio for each stratum is weighted by the number of people in each stratum. We do this for all 10,000 samples, and then generate a mean and confidence intervals for the calibrated relative risk.

```{r pooling hrs}
#matrix of risk ratios at t1, one column per stratum

rr_t1<-matrix(risk_ratio_t1, ncol=4)
rr_t2<-matrix(risk_ratio_t2, ncol=4)

#matrix of ns in latch, one column per stratum
strata_n_df<- as.data.frame(t(strata_n))
ns <- as.matrix(strata_n_df %>% slice(rep(1, each = nrep)))
#create a vector of totals
n_tot<- rowSums(ns)

## calculate n x rr_mtrx (note that this is NOT matrix multiplication)
#t1
rn_t1 <- rowSums(rr_t1*ns)
rr_tot_t1 <- rn_t1/n_tot
cis_t1<- round(quantile(rr_tot_t1, probs = c(0.025, 0.975)),3)
est_t1<- round(mean(rr_tot_t1),3)

#t2
rn_t2 <- rowSums(rr_t2*ns)
rr_tot_t2 <- rn_t2/n_tot
cis_t2<- round(quantile(rr_tot_t2, probs = c(0.025, 0.975)),3)
est_t2<- round(mean(rr_tot_t2),3)
```



#### Pooled risk ratio at T1 + T2

```{r t2 pooled results,results="asis"}

cat(paste(" \n \n The estimated calibrated risk ratio is ", est_t2, " (95% confidence interval ", cis_t2[1], "-", cis_t2[2], ")", sep=""))

ggplot(data=as.data.frame(rr_tot_t2))+
  geom_histogram(aes(x=rr_tot_t2), fill="#80B1D3")+
  theme_bw()+
  labs(title="Simulated risk ratio, T1+T2")+ 
  xlab("Risk ratio")


```

### Absolute risk 

We can use the new, calibrated risk ratio to estimate the number needed to treat (NNT) from the absolute risk ratio (ARR).

$$ARR = 1/(\lambda_u - \lambda_u \times RR)$$

$$ \text{NNT} = 1/\text{ARR} $$

```{r absolute risks}
event_rate<- sum(strata_n_events_t3)/sum(strata_fu_t3)*365.25

arr <- event_rate- event_rate*est_t2

nnt <- 1/arr

cat(paste(" \n \n The number needed to treat is ",nnt, sep=""))

```

### Appendix

We can pool the risk ratio at T1, to find the calibrated relative risk after 18 months.

#### Pooled risk ratio at T1

```{r t1 pooled results,results="asis"}
cat(paste(" \n \nThe estimated calibrated risk ratio is ", est_t1, " (95% confidence interval ", cis_t1[1], "-", cis_t1[2], ")", sep=""))

ggplot(data=as.data.frame(rr_tot_t1))+
  geom_histogram(aes(x=rr_tot_t1), fill="#FCCDE5")+
  theme_bw()+
  labs(title="Simulated risk ratio, T1")+ 
  xlab("Risk ratio")

```




