---
title: "Mock-up of trial calibration"
author: "Alice Hosking"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(flexsurv)
library(survival)
library(SurvRegCensCov)
library(muhaz)
library(kableExtra)
library(ggsurvfit)
library(survminer)
library(rmeta)

set.seed(42)

knitr::opts_chunk$set(message = F, warning=F, echo=F)
```

## Introduction

This document contains sample R code and results for trial calibration based on strata. The intention is to calibrate TRIDENT to LATCH. The TRIDENT trial result showed that there appeared to be no difference in the primary outcome until approximately 18 months, from which point the cumulative incidence function for the primary outcome appears to diverge between control and treatment arms. We will therefore calibrate at two time points, 18 months (t1) and end of follow up (approx 6 years, t1+t2=t3).

In this document, we use real LATCH data and made-up hazard ratios, to demonstrate the method.

The method will be as follows:

* Inspect the distribution of ages in TRIDENT and LATCH to choose sensible stratification ages.
* Split TRIDENT into strata of age and sex (eg <65 years female, <65 years male, over 65 years female,  over 65 years male)
* Obtain hazard ratios for TRIDENT for each stratum at 18 months and 4.5 years. 
* Split LATCH into the same age- and sex-based strata.For each strata, fit a parametric survival model to obtain rates and confidence intervals of the primary outcome at t1 and t2.
* Use these rates, and the hazard ratios from TRIDENT, to model absolute rates in untreated and treated in each stratum.
* The relative risk for each stratum is then (risk untreated/risk treated). We combind relative risks and use simulation to obtain confidence intervals.

## Data tidying

First we filter LATCH to only those who survived 7 days, format dates correctly, set censoring dates, and split into strata (under 65 and 65 and over, male and female). (Code not shown.)

```{r read in data, echo=FALSE, message=FALSE}
latch_for_tridentx<- read_csv("/Users/alicehosking/Documents/REA3 clinical fellow job/Treatment effect in trials/LATCH cohort survived 7 days.csv")
setDT(latch_for_tridentx)
latch_for_trident<- latch_for_tridentx[,c("survival_time","censor_status_death","i_date","Gender_t", "date_of_death.y","cause_of_death","first_mi","first_tia","first_stroke","first_dvt","first_pe", "first_rec_ich","pre_rankin", "age_index")] 
#some dates are stored as excel formatted dates. 
#for this analysis, we care about date of first stroke (which is ischaemic stroke) and date of recurrent ICH
latch_for_trident[, first_rec_ich:= as.Date(first_rec_ich, origin = "1899-12-30")][, first_stroke:= as.Date(first_stroke, origin = "1899-12-30")]

#time to first event
latch_for_trident[, time_to_ich:= difftime(first_rec_ich,i_date,  units="days")][,time_to_isch_str:=difftime(first_stroke ,i_date,  units="days")][, time_to_all_stroke:= pmin(time_to_ich, time_to_isch_str, na.rm=T)][,date_of_death_trid:= as.Date(date_of_death.y, origin = "1899-12-30")] [, time_to_death:= difftime(date_of_death_trid,i_date,  units="days")]

# if no first event, time to end of death/ end study
latch_for_trident[, surv_time_all_stroke:=fifelse(is.na(time_to_all_stroke)==T&is.na(time_to_death)==T, difftime("2022-12-31", i_date, units="days"), fifelse(is.na(time_to_all_stroke)==T&is.na(time_to_death)==F, time_to_death, time_to_all_stroke))]

#censor status and survival time at 1.5 years, 4.5 years and all time
latch_for_trident[, censor_status_all_str_all_time:= fifelse(is.na(time_to_all_stroke)==T,0,1)][,censor_all_str_t1:= fifelse(time_to_all_stroke<548,1,0)][ ,censor_all_str_t1:=fifelse(is.na(time_to_all_stroke)==T,0,censor_all_str_t1)][, survival_stroke_t1:= fifelse(censor_all_str_t1==1, time_to_all_stroke, pmin(surv_time_all_stroke, time_to_death, na.rm = T))][, survival_stroke_t1:= pmin(548,survival_stroke_t1)][,censor_all_str_t3:= fifelse(time_to_all_stroke<6*365.25,1,0)][ ,censor_all_str_t3:=fifelse(is.na(time_to_all_stroke)==T,0,censor_all_str_t3)][, survival_stroke_t3:= fifelse(censor_all_str_t3==1, time_to_all_stroke, pmin(surv_time_all_stroke, time_to_death, na.rm = T))][,survival_stroke_t3:=pmin(round(6*365.25,0),survival_stroke_t3)]

strata<- list(
latch_under_65_m<- latch_for_trident[age_index<65&Gender_t=="Male"],
latch_under_65_f<- latch_for_trident[age_index<65&Gender_t=="Female"],
latch_over_65_m<- latch_for_trident[age_index>64&Gender_t=="Male"],
latch_over_65_f<- latch_for_trident[age_index>64& Gender_t=="Female"]
  
)

strata_n<- c()
strata_n_events_t1<- c()
strata_n_events_t3<- c()
strata_fu_t1 <- c()
strata_fu_t3<- c()

for(i in 1:length(strata)){
  strata_n[i] <-length(strata[[i]]$i_date) 
  strata_n_events_t1[i]<- sum(strata[[i]]$censor_all_str_t1)
  strata_n_events_t3[i]<- sum(strata[[i]]$censor_all_str_t3)
  strata_fu_t1[i] <- sum(strata[[i]]$survival_stroke_t1)
  strata_fu_t3[i]<- sum(strata[[i]]$survival_stroke_t3)

}

```

## Simple event rates {.tabset}

Assuming exponential event rate across t1 and t2, we can calculate the event rate, the standard error, and simulate a normally distributed sample. We plot the sample and show the event rates for each stratum in a table.

### Plot of sample

```{r simple event rates}
#event rate at t1 and t2
log_rate_est <- log((strata_n_events_t3)/ strata_fu_t3)
# Calculate SE of the log-rate for untreated
log_rate_se <- 1 / sqrt(strata_n_events_t3)
## obtain and plot samples from rate in untreated
## Set number of samples with nrep for use throughout.
nrep<-1000
untreated_rates<- as.data.frame(cbind(log_rate_est, log_rate_se))
setDT(untreated_rates)

log_lambda_u<- untreated_rates[,rnorm(nrep, mean=log_rate_est, sd=log_rate_se), by=.I]

#plot to confirm normality
ggplot(data=log_lambda_u)+
  geom_histogram(aes(x=V1), fill="#FCCDE5")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated log rates for strata 1-4")+ 
  xlab("")
```

### Log of event rates for each stratum

```{r table of rates}
#print table of rates
cbind(c("Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over"), round(untreated_rates,3))%>%
  kable()%>%
  kable_styling()

```


## Hazard ratios for each stratum {.tabset}

For this mock-up we estimate hazard ratios for each time period. For the real analysis, the strata may be different; this is for demonstration only. We assume HR are independent; and make a guess at the HR for young men. We then split into two time periods: up to 18 months, where the HR are approximately 1; and beyond 18 months. We simulate 1000 samples assuming a log normal distribution, for t1 and t2.

### Table of mock hazard ratios

```{r mock hazard ratios - whole trial}
# overall HR. These come from the subgroup analysis of the trial
#create vector for each with HR, LCI, UCI
male_hr<- c(0.71,0.42,1.15)
female_hr<- c(0.44,0.21,0.94)
f_m_interaction <- female_hr/male_hr

under_60_hr<- c(0.76,0.44,1.33)
over_60_hr<- c(0.61,0.41,0.92)
old_young_interaction<- over_60_hr/under_60_hr

#assume independence; can multiply to get HR for subgroups
#need to make assumption on HR for young men

hrs_t2<- as.data.frame(rbind(
hr_young_men<- c(0.78, 0.46, 1.35),
hr_young_women <- hr_young_men*f_m_interaction,
hr_old_men<- hr_young_men*old_young_interaction,
hr_old_women <- hr_young_men*f_m_interaction*old_young_interaction
))
# we need to split into two time strata; the first 18 months and beyond 18 months. For the first 18 months, the HR between treatment and control overall is approximately 1; beyond this the HR widens. for the first 18 months, recentre HR and CI to 1ish. For beyond 18 months, use existing HR
hrs_t1<- as.data.frame(rbind(
hr_young_men_t1<- hr_young_men+0.22,
hr_young_women_t1 <- hr_young_women+0.52,
hr_old_men_t1 <- hr_old_men+0.48,
hr_old_women_t1 <- hr_old_women + 0.62
))

colnames(hrs_t1) <- c("HR", "l_ci", "u_ci")
rownames(hrs_t1) <- c(" Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")
colnames(hrs_t2) <- c("HR", "l_ci", "u_ci")
rownames(hrs_t2) <- c(" Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")

kable(hrs_t1)%>% kable_styling()
kable(hrs_t2)%>% kable_styling()
```

### Histogram of simulated log hazard ratios for strata 1-4, T1

```{r simulate n hrs, warnings=F}
#obtain log of hrs and CI at t1 and t2. Assume log normally distributed, and find log standard error
setDT(hrs_t1)
setDT(hrs_t2)
hrs_t1[, log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]
hrs_t2[, log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]

#simulate log hr for nrep samples.
#create simulated hr for all strata
log_hrs_t1_per_strata <- hrs_t1[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]
#format as matrix
log_hrs_t1_per_strata_mat <- as.data.frame(matrix(log_hrs_t1_per_strata$V1, ncol=4))
#plot to confirm normality
ggplot(data=log_hrs_t1_per_strata)+
  geom_histogram(aes(x=V1), fill="#FCCDE5")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated log hazard ratios for strata 1-4, T1")+ 
  xlab("Log of hazard ratio")
```

### Histogram of simulated hazard ratios for strata 1-4, T1 

```{r hist hr t1 exponentiated}
ggplot(data=log_hrs_t1_per_strata)+
  geom_histogram(aes(x=exp(V1)), fill="#FCCDE5")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated hazard ratios for strata 1-4, T1")+ 
  xlab("Hazard ratio")
```

### Histogram of simulated log hazard ratios for strata 1-4, T2

```{r t2 hr hists}
#repeat for t2 hazard ratios
nrep<-1000
#create simulated hr for all strata
log_hrs_t2_per_strata <- hrs_t2[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]
#format as matrix
log_hrs_t2_per_strata_mat <- as.data.frame(matrix(log_hrs_t2_per_strata$V1, ncol=4))
#plot to confirm normality
ggplot(data=log_hrs_t2_per_strata)+
  geom_histogram(aes(x=V1), fill="#80B1D3")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated log hazard ratios for strata 1-4, T2")+ 
  xlab("Log of hazard ratio")
```

### Histogram of simulated hazard ratios for strata 1-4, T2 

```{r t2 hr exponentiated}
ggplot(data=log_hrs_t2_per_strata)+
  geom_histogram(aes(x=exp(V1)), fill="#80B1D3")+
  theme_bw()+
  facet_wrap(vars(I))+
  labs(title="Simulated hazard ratios for strata 1-4, T2")+ 
  xlab("Hazard ratio")
```

## Calibrated rates and CIs for each stratum {.tabset}

For each age/sex stratum and for each time period, we combine the rate in the untreated and the hazard ratio to obtain a risk ratio.

The basic formula for an exponential distribution are as follows:

lambda_u - rate in untreated
t1=length in days of first time period
t2=length in days of second time period

Multiply untreated rates by  hazard ratios to get treated event rates:-
lamba_t1 = lambda_u * hr1
lamba_t2 = lambda_u * hr2

We then transform these to risks for each stratum
Risk_untreated = 1 - exp(-lambda_u*(t1+t2))
Risk_treated = 1 - exp(-(lambda_t1 * t1 + lambda_t2*t2))

We have already simulated 1000 samples for each hazard ratio. We simulate 1000 samples for event rates, which gives 1000 risk ratios.

```{r risks treated and untreated 2}
#set t1 and t2 in days
t1<-548
t2<- 4.5*365.25

#lambda from parametric model above
#lambda_u <- exp(z_lambda) 
log_lambda_u_sim<- log_lambda_u$V1

#for rate in treated, log lambda_t = log lambda_u + log hr
log_lambda_t1_sim <- log_lambda_u_sim+log_hrs_t1_per_strata$V1
log_lambda_t2_sim <- log_lambda_u_sim+log_hrs_t2_per_strata$V1

#exponentiate to get rates
lambda_u_sim <- exp(log_lambda_u_sim)
lambda_t1_sim<- exp(log_lambda_t1_sim)
lambda_t2_sim<- exp(log_lambda_t2_sim)

#transform to risks
risk_u1 <- 1-exp(-lambda_u_sim*t1)
risk_t1 <- 1-exp(-lambda_t1_sim*t1)
risk_u2 <- 1-exp(-lambda_u_sim*(t1+t2))
risk_t2 <- 1-exp(-(lambda_t1_sim*t1+lambda_t2_sim*t2))
risk_ratio_t1 <- risk_t1/risk_u1
risk_ratio_t2 <- risk_t2/risk_u2

risk_ratio_for_plot_t1 <- as.data.frame(cbind(c(rep(1,nrep), rep(2, nrep), rep(3, nrep), rep(4, nrep)), risk_ratio_t1))
risk_ratio_for_plot_t2 <- as.data.frame(cbind(c(rep(1,nrep), rep(2, nrep), rep(3, nrep), rep(4, nrep)), risk_ratio_t2))

rr_means_t1<- risk_ratio_for_plot_t1%>%
  group_by(V1)%>%
  summarize(means_t1=round(mean(risk_ratio_t1),3))

rr_means_t2<- risk_ratio_for_plot_t2%>%
  group_by(V1)%>%
  summarize(means_t2=round(mean(risk_ratio_t2),3))

```

### Simulated risk ratios for strata 1-4, T1 

```{r risk plot t1}
ggplot(data=risk_ratio_for_plot_t1)+
  geom_histogram(aes(x=risk_ratio_t1), fill="#FCCDE5")+
  geom_text(data=rr_means_t1, aes(x=1, y=150, label= paste("Mean=",means_t1)))+
  theme_bw()+
  facet_wrap(vars(V1))+
  labs(title="Simulated risk ratios for strata 1-4, T1")+ 
  xlab("Risk ratio")
```

### Simulated risk ratios for strata 1-4, T1+T2

```{r risk plot t2}
ggplot(data=risk_ratio_for_plot_t2)+
  geom_histogram(aes(x=risk_ratio_t2), fill="#80B1D3")+
  geom_text(data=rr_means_t2, aes(x=1, y=150, label= paste("Mean=",means_t2)))+
  theme_bw()+
  facet_wrap(vars(V1))+
  labs(title="Simulated risk ratios for strata 1-4, T1+T2")+ 
  xlab("Risk ratio")

```

## Pooling risk ratios {.tabset}

The final step is to pool the risk ratios for each stratum.

```{r pooling hrs}
#matrix of risk ratios at t1, one column per stratum

rr_t1<-matrix(risk_ratio_t1, ncol=4)
rr_t2<-matrix(risk_ratio_t2, ncol=4)

#matrix of ns in latch, one column per stratum
strata_n_df<- as.data.frame(t(strata_n))
ns <- as.matrix(strata_n_df %>% slice(rep(1, each = 1000)))
#create a vector of totals
n_tot<- rowSums(ns)

## calculate n x rr_mtrx (note that this is NOT matrix multiplication)
#t1
rn_t1 <- rowSums(rr_t1*ns)
rr_tot_t1 <- rn_t1/n_tot
cis_t1<- round(quantile(rr_tot_t1, probs = c(0.025, 0.975)),3)
est_t1<- round(mean(rr_tot_t1),3)

#t2
rn_t2 <- rowSums(rr_t2*ns)
rr_tot_t2 <- rn_t2/n_tot
cis_t2<- round(quantile(rr_tot_t2, probs = c(0.025, 0.975)),3)
est_t2<- round(mean(rr_tot_t2),3)
```

### Pooled risk ratio at T1

```{r t1 pooled results,results="asis"}
cat(paste(" \n \nThe estimated calibrated risk ratio is ", est_t1, " (95% confidence interval ", cis_t1[1], "-", cis_t1[2], ")", sep=""))

ggplot(data=as.data.frame(rr_tot_t1))+
  geom_histogram(aes(x=rr_tot_t1), fill="#80B1D3")+
  theme_bw()+
  labs(title="Simulated risk ratio, T1")+ 
  xlab("Risk ratio")



```

### Pooled risk ratio at T1 + T2

```{r t2 pooled results,results="asis"}

cat( paste(" \n \n The estimated calibrated risk ratio is ", est_t2, " (95% confidence interval ", cis_t2[1], "-", cis_t2[2], ")", sep=""))

ggplot(data=as.data.frame(rr_tot_t2))+
  geom_histogram(aes(x=rr_tot_t2), fill="#FCCDE5")+
  theme_bw()+
  labs(title="Simulated risk ratio, T1+T2")+ 
  xlab("Risk ratio")


```





