---
title: "Calibration of the TRIDENT trial to a population-based
cohort in Scotland."
author: "Dr Alice Hosking, Dr Katie Harris, Prof Rustam Al-Shahi Salman, Prof William Whiteley, Prof David McAllister"
email: "alice.hosking@ed.ac.uk"
date: "`r Sys.Date()`"
output:
  html_document:
   toc: true
   number_sections: true
   theme: united
   
  # word_document:
  #  toc: true
bibliography: calibration.bib
csl: "american-medical-association-10th-edition.csl"

---


```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(flexsurv)
library(survival)
library(SurvRegCensCov)
library(muhaz)
library(ggsurvfit)
library(survminer)
library(rmeta)
library(flextable)
library(patchwork)
library(readxl)
library(RColorBrewer)

set.seed(42)

knitr::opts_chunk$set(message = F, warning=F, echo=F)
times_for_tables<- c(0, 90,365, 365*5)
set_flextable_defaults(
  font.size = 10, theme_fun = theme_vanilla,
  padding = 6,
  background.color = "#FFFFFF")


#This function calculates the cumulative incidence accounting for competing risks.

MakeCumulative2 <- function (arm, rate.main.t1, rate.main.t2, rate.compete.t1, rate.compete.t2, t1_times, t2_times){

# 
# arm="u"
#                        rate.main.t1 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)]
#                        rate.main.t2 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)]
#                        rate.compete.t1 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2])
#                        rate.compete.t2 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2])

#combine t1 and t2 to get total time
all_times <- c(t1_times, t2_times)

#define empty matrices
lambda_new <- matrix(NA, ncol = 2, nrow = length(all_times)*nrep)
p_new <- matrix(NA, ncol = 3, nrow = length(all_times)*nrep)
cumslam_new <- NA

#fill rate matrix - one column each for rate of event of interest (rec stroke) and competing risk (death)
#column 1 is rec stroke
lambda_new_t1  <- rep(rate.main.t1, length(t1_times))
lambda_new_t2<- rep(rate.main.t2, length(t2_times))
lambda_new [1:(length(t1_times)*nrep),1] <- lambda_new_t1
lambda_new[((length(t1_times)*nrep)+1):(length(all_times)*nrep),1]<- lambda_new_t2

#column 2 is competing event
lambda_new_t1_comp  <- rep(rate.compete.t1$V1, length(t1_times))
lambda_new_t2_comp <- rep(rate.compete.t2$V1, length(t2_times))
lambda_new [1:(length(t1_times)*nrep),2] <- lambda_new_t1_comp
lambda_new[((length(t1_times)*nrep)+1):(length(all_times)*nrep),2]<-lambda_new_t2_comp


# lambda_new_t1 [,2] <- rep(rate.compete.t1$V1, length(all_times))

#sum these to get lambda event + lambda comp
slam_new_1 <- rowSums(lambda_new[1:(length(t1_times)*nrep),1:2]) 
slam_new_2 <- rowSums(lambda_new[((length(t1_times)*nrep)+1):(length(all_times)*nrep),1:2]) 
slam_new <- c(slam_new_1, slam_new_2)

#cumulative rates. First, for both outcomes
cumslam_new_1 <- 1 - exp(-(slam_new_1*rep(t1_times, each=nrep)))
# need to adjust start point at end of t1: first day of t2 - last day of t1. 
adj_cum_x <- exp(-slam_new_2[1:nrep]*(rep(t2_times[1],nrep))) - exp(-slam_new_1[((length(t1_times)-1)*nrep+1):(length(t1_times)*nrep)]* rep(max(t1_times), nrep))

#calculate for t2
cumslam_new_2<- 1-exp(-slam_new_2*rep(t2_times, each=nrep)) + rep(adj_cum_x, length(t2))
#join together  
cumslam_new <- c(cumslam_new_1, cumslam_new_2)

# Calculate cumulative incidence for each separate outcome. Because of the two times, we need to do each separately (rather than the nice simple proportions in DM's original code)
#for t1, calculate cum incidence of outcome j
cumslam_outcome_j_t1 <- 1 - exp(-(lambda_new_t1*rep(t1_times, each=nrep)))
# need to adjust start point at 2t2:  first day of t2 - last day of t1 
# adj_cum_j_x <- exp(-lambda_new_t2[1:nrep]*(rep(t2_times[1],nrep))) - exp(-(lambda_new_t1[(max(t1_times)*(nrep-1)+1): (max(t1_times)*nrep)]*rep(max(t1_times), nrep)))
# # 
day_1_t2 <-exp(-lambda_new_t2[1:nrep]*(rep(t2_times[1],nrep)))
day_n_t1 <- exp(-(lambda_new_t1[(((length(t1_times)-1)*nrep)+1): (length(t1_times)*nrep)]*rep(max(t1_times), nrep)))

adj_cum_j_x <- day_1_t2 - day_n_t1 

cumslam_outcome_j_t2<- 1-exp(-lambda_new_t2*rep(t2_times, each=nrep)) + rep(adj_cum_j_x, length(t2_times))
p_new[,1] <- c(cumslam_outcome_j_t1,cumslam_outcome_j_t2 )

#do same for competing risk
cumslam_outcome_comp_t1 <- 1 - exp(-(lambda_new_t1_comp*rep(t1_times, each=nrep)))

day_1_t2_comp <-exp(-lambda_new_t2_comp[1:nrep]*(rep(t2_times[1],nrep)))
day_n_t1_comp <- exp(-(lambda_new_t1_comp[(((length(t1_times)-1)*nrep)+1): (length(t1_times)*nrep)]*rep(max(t1_times), nrep)))

adj_cum_comp_x <- day_1_t2_comp - day_n_t1_comp 

cumslam_outcome_comp_t2<- 1-exp(-lambda_new_t2_comp*rep(t2_times, each=nrep)) + rep(adj_cum_comp_x, length(t2_times))


p_new[,2] <-  c(cumslam_outcome_comp_t1, cumslam_outcome_comp_t2)

p_new[,3] <- cumslam_new

p_new <- as.data.frame(p_new)
}


#For each output, will need mean of each day, and ci of each day. This function calculates this.
get_CIs <- function(x, times){
df<- setDT(as.data.frame(x))
df_new <- data.frame(matrix(nrow=length(times), ncol=9))
df[, index:=rep(times, each=nrep)]

#Group by index then work out quantiles. Use data.table
df[,mean_main:= mean(V1,na.rm=T), by=index][,l_ci_main:=quantile(V1, probs=0.025, na.rm=T),by=index][, u_ci_main:=quantile(V1, probs=0.975,na.rm=T), by=index]
df[,mean_comp:= mean(V2,na.rm=T), by=index][,l_ci_comp:=quantile(V2, probs=0.025,na.rm=T),by=index][, u_ci_comp:=quantile(V2, probs=0.975,na.rm=T), by=index]
df[,mean_both:= mean(V3,na.rm=T), by=index][,l_ci_both:=quantile(V3, probs=0.025,na.rm=T),by=index][, u_ci_both:=quantile(V3, probs=0.975,na.rm=T), by=index]

#select only relevant columns and every ntimes row
df_new <- df[seq(1, nrow(df),nrep), ]
df_new
}



```

## Introduction

### Purpose and scope of this document

This is a brief report of the results of calibration of the TRIDENT trial to a population-based inception cohort. 
The aim of this study is to calibrate the effect of a low-dose triple pill antihypertensive medication in people with ICH from the TRIDENT trial to a population-based cohort in Lothian, Scotland.

The outcomes are:

* A measure of relative effect, the risk ratio, calibrated to the Lothian population.
* A measure of absolute effect, the number needed to treat, calibrated to the Lothian population.
* A measure of absolute effect, the absolute risk reduction in the presence of competing risk of death, calibrated to the Lothian population.

## Methods

### Time period definitions

We calibrate at two time points, 18 months ($t_1$) and end of follow up (approx 6 years, $t_2$). The time period from start of follow-up to $t_1$ is $\tau_1$. The time period from $t_1$ to $t_2$ is $\tau_2$.

### Strata for analysis

Following inspection of the overlap between LATCH and TRIDENT, the split is as follows:
TRIDENT is split into four strata: Male <65 years, Female <65 years, Male 65 years and over, Female 65 years and over.
LATCH is split into eight strata: Male <65 years, Female <65 years, Male 65-75 years, Female 65-75 years,Male 75-85 years, Female  75-85 years,Male 85 years and over, Female 85 years and over.

### Data preparation

First we filter LATCH to only those who survived 30 days, format dates correctly, set censoring dates, and split into strata (under 65 and 65 and over, male and female). For this study, we censor LATCH follow up at the maximum TRIDENT follow-up time of six years.

### Event rates in LATCH

Assuming exponential event rate across $\tau_1$ and $\tau_2$, we can calculate the event rate, the standard error, and simulate a normally distributed sample.@Greenland2004 Throughout, we simulate with 1000 samples. We plot the sample and show the event rates for each stratum in a table.

The event rate is defined as:

$$\lambda_u=\text{rate in untreated}$$
We define $n_s$ as the number of people in each stratum, and $m_s$ as the number of events per stratum.

$$\lambda_u=\frac{m_s}{\text{Total follow up time for stratum}}$$.

As we are assuming an exponential distribution, we can calculate the standard error directly:

$$ ln(SE)= \frac{1}{\sqrt{m_s}}$$

We then simulate a normally distributed sample for the event rates.

### Calibrated RR and CIs for each stratum

For each age/sex stratum and for each time period, we combine the rate in the untreated and the hazard ratio from the trial to obtain a risk ratio.

Recall:

$$\lambda_u=\text{rate in untreated}$$


Multiply untreated rates by hazard ratios ($HR$) to get treated event rates in each time period:
$$\lambda_{treat1} = \lambda_u \times HR_1$$
$$\lambda_{treat2} = \lambda_u \times HR_2$$

We then transform these to risks for each stratum:
$$R_{untreated} = 1 - exp{(-\lambda_u(\tau_1+\tau_2)})$$ 

and 

$$R_{treated} = 1 - exp{(-(\lambda_{treat1} \tau_1 + \lambda_{treat2} \tau_2)})$$

The risk ratio at the end of the trial for each stratum is then:

$$ RR_S= \frac{R_{treated}}{R_{untreated}}$$
  
We simulate 1000 samples for each hazard ratio and event rate, to generate 1000 risk ratios.

### Pooled risk ratio

The final step is to pool the risk ratio for each stratum. We take a weighted mean, where the risk ratio for each stratum is weighted by the number of people in each stratum. 

$$RR_{overall}= \sum_{s} RR_s n_s $$
  
  We do this for all samples, then generate a mean and confidence intervals for the calibrated relative risk.

### Number needed to treat

We can use the new, calibrated risk ratio to estimate the number needed to treat (NNT) in the LATCH cohort from the absolute risk ratio (ARR). We use 1000 samples of the risk ratio, and generate confidence intervals.

$$ARR = 1/(\lambda_u - \lambda_u \times RR_{overall})$$
  
  $$ \text{NNT} = 1/\text{ARR} $$

### Absolute risk in the presence of competing risks

People with intracerebral haemorrhage have a high mortality, and this might have a significant effect on the absolute efficacy of treatment for particular groups. We now take into account competing risk of death from a non-stroke cause.

Assuming exponential event rate across $\tau_1$ and $\tau_2$, we calculate the event rate, the standard error, and simulate a normally distributed sample. 

The competing risk of death might not be independent of treatment with the study drug, since we would expect antihypertensives to reduce death from other cardiovascular causes. We therefore perform two analysis; first, with no effect on competing risk of the study drug; and second, we include a hazard ratio for the competing risk between the placebo arm and treatment arm, for $\text{t}_1$ and $\text{t}_2$ and for individual strata.


### Calculate event rate accounting for competing risk

The proportion event-free at each time point, taking into account competing risk of death, is:
  
  $$\text{proportion event-free} = 1 - exp{(-\lambda_i t})-exp{(-\lambda_j  t}) $$
      
The cumulative incidence of the outcome event $i$ is then:
      
  $$ \text{cumulative incidence} = \text{proportion event-free} \times \lambda_i / (\lambda_i + \lambda_j)$$
        
However, if the rate changes over time, the above formula doesn't work; instead, we calculate the cumulative incidence for each event directly. For two time periods with different hazard ratios, the cumulative incidence needs adjusting at the point the hazard ratio changes, by a constant $c$. For two time periods, $\tau_1 = (0,\text{t}_1)$ and $\tau_2 = (\text{t}_1+1, \text{t}_2)$, with rates $\lambda_{i1}$ and $\lambda_{i2}$:

$$ c =-exp{(-\lambda_{i1} \text{t}_1}) - exp{(-\lambda_{i2}(\text{t}_1+1))} $$
The cumulative incidence for times beyond $\text{t}_1+1$ is:

$$\text{proportion event-free} = 1 - exp{(-\lambda_i2 t})-exp{(-\lambda_j  t}) +c $$
As with the risk ratio analysis, I account for two time periods with different rates of outcome event.

We assume that:

* Both the risk of recurrent stroke and the competing risk of death are adequately modelled as exponential distributions.
* People in LATCH are commenced on the study drug at 30 days post-stroke.

## Results

### LATCH data {.tabset}

#### Event numbers at T1 and T2

The table shows the number of recurrent stroke events and non-stroke death events in LATCH for eight strata.

```{r read in data, echo=FALSE, message=FALSE}
latch_for_tridentx<- read_csv("/Users/alicehosking/Documents/REA3 clinical fellow job/Treatment effect in trials/LATCH cohort survived 30 days.csv")
setDT(latch_for_tridentx)
latch_for_trident<- latch_for_tridentx[,c("survival_time","censor_status_death","i_date","Gender_t", "date_of_death.y","cause_of_death","first_mi","first_tia","first_stroke","first_dvt","first_pe", "first_rec_ich","pre_rankin", "age_index")] 
#some dates are stored as excel formatted dates. 
#for this analysis, we care about date of first stroke (which is ischaemic stroke) and date of recurrent ICH
latch_for_trident[, first_rec_ich:= as.Date(first_rec_ich, origin = "1899-12-30")][, first_stroke:= as.Date(first_stroke, origin = "1899-12-30")]

#time to first event
latch_for_trident[, time_to_ich:= difftime(first_rec_ich,i_date,  units="days")][,time_to_isch_str:=difftime(first_stroke ,i_date,  units="days")][, time_to_all_stroke:= pmin(time_to_ich, time_to_isch_str, na.rm=T)][,date_of_death_trid:= as.Date(date_of_death.y, origin = "1899-12-30")] [, time_to_death:= difftime(date_of_death_trid,i_date,  units="days")]

# if no first event, time to end of death/ end study
latch_for_trident[, surv_time_all_stroke:=fifelse(is.na(time_to_all_stroke)==T&is.na(time_to_death)==T, difftime("2022-12-31", i_date, units="days"), fifelse(is.na(time_to_all_stroke)==T&is.na(time_to_death)==F, time_to_death, time_to_all_stroke))]

#censor status and survival time at 1.5 years, 4.5 years and all time
latch_for_trident[, censor_status_all_str_all_time:= fifelse(is.na(time_to_all_stroke)==T,0,1)][,censor_all_str_t1:= fifelse(time_to_all_stroke<548,1,0)][ ,censor_all_str_t1:=fifelse(is.na(time_to_all_stroke)==T,0,censor_all_str_t1)][, survival_stroke_t1:= fifelse(censor_all_str_t1==1, time_to_all_stroke, pmin(surv_time_all_stroke, time_to_death, na.rm = T))][, survival_stroke_t1:= pmin(548,survival_stroke_t1)][,censor_all_str_t2:= fifelse(time_to_all_stroke<6*365.25,1,0)][ ,censor_all_str_t2:=fifelse(is.na(time_to_all_stroke)==T,0,censor_all_str_t2)][, survival_stroke_t2:= fifelse(censor_all_str_t2==1, time_to_all_stroke, pmin(surv_time_all_stroke, time_to_death, na.rm = T))][,survival_stroke_t2:=pmin(round(6*365.25,0),survival_stroke_t2)]

# here we also define the survival times for competing risk of death, which we use in the extension below
#create variable to say if has competing event
#calculate time to competing event
#censor: no death, or stroke before death from other cause.
#provide interim count at t1

#censor if had stroke/didn't die
latch_for_trident[, comp_event_n:= fifelse(is.na(time_to_all_stroke)==T & is.na(date_of_death.y)==F,1,0) ]
#time to event is: time to stroke (for people who had stroke); time to death (no stroke and died from other cause);  end of follow up. Coded using fcase.
latch_for_trident[, time_to_comp_event := fcase(is.na(time_to_all_stroke)==F, time_to_all_stroke, is.na(date_of_death.y)==F& is.na(time_to_all_stroke)==T,difftime(date_of_death_trid, i_date, units="days"),default=difftime("2022-12-31", i_date, units="days"))]
#provide interim count at t1
latch_for_trident[, comp_event_n_t1 := fifelse(time_to_comp_event < 548, 1,0)][, time_to_comp_event_t1:= pmin(548,time_to_comp_event)][, comp_event_n_t2 := fifelse(time_to_comp_event < 6*365.25, 1,0)][,time_to_comp_event_t2 := pmin(round(6*365.25),time_to_comp_event)]

strata<- list(
latch_under_65_m<- latch_for_trident[age_index<65&Gender_t=="Male"],
latch_under_65_f<- latch_for_trident[age_index<65&Gender_t=="Female"],
latch_over_65_m<- latch_for_trident[age_index>64&Gender_t=="Male"],
latch_over_65_f<- latch_for_trident[age_index>64& Gender_t=="Female"]
)

strata_n<- c()
strata_n_events_t1<- c()
strata_n_events_t2<- c()
strata_fu_t1 <- c()
strata_fu_t2<- c()
strata_comp_n_events_t1 <- c()
strata_comp_n_events_t2 <- c()
strata_comp_fu_t1<- c()
strata_comp_fu_t2 <- c()

for(i in 1:length(strata)){
strata_n[i] <-length(strata[[i]]$i_date) 
strata_n_events_t1[i]<- sum(strata[[i]]$censor_all_str_t1)
strata_n_events_t2[i]<- sum(strata[[i]]$censor_all_str_t2)
strata_fu_t1[i] <- sum(strata[[i]]$survival_stroke_t1)
strata_fu_t2[i]<- sum(strata[[i]]$survival_stroke_t2)

strata_comp_n_events_t1[i]<- sum(strata[[i]]$comp_event_n_t1)
strata_comp_n_events_t2[i]<- sum(strata[[i]]$comp_event_n_t2)
strata_comp_fu_t1[i] <- sum(strata[[i]]$time_to_comp_event_t1)
strata_comp_fu_t2[i]<- sum(strata[[i]]$time_to_comp_event_t2)


}
strata_vars<- list(strata_n=strata_n, strata_n_events_t1=strata_n_events_t1, strata_n_events_t2=strata_n_events_t2, strata_fu_t1=strata_fu_t1, strata_fu_t2=strata_fu_t2, strata_comp_n_events_t1=strata_comp_n_events_t1, strata_comp_n_events_t2=strata_comp_n_events_t2, strata_comp_fu_t1=strata_comp_fu_t1, strata_comp_fu_t2=strata_comp_fu_t2)
saveRDS(strata_vars, "RDS files/n per strata, 65yrs.rds")

#alternative strata for competing risk
strata_y<- list(
latch_under_65_m<- latch_for_trident[age_index<65&Gender_t=="Male"],
latch_under_65_f<- latch_for_trident[age_index<65&Gender_t=="Female"],
latch_65_75_m<- latch_for_trident[age_index>64&age_index<75&Gender_t=="Male"],
latch_65_75_f<- latch_for_trident[age_index>64&age_index<75& Gender_t=="Female"],
latch_75_85_m<- latch_for_trident[age_index>74&age_index<85&Gender_t=="Male"],
latch_75_85_f<- latch_for_trident[age_index>74&age_index<85& Gender_t=="Female"],
latch_over_85_m<- latch_for_trident[age_index>84&Gender_t=="Male"],
latch_over_85_f<- latch_for_trident[age_index>84& Gender_t=="Female"]
)

strata_y_n<- c()
strata_y_n_events_t1<- c()
strata_y_n_events_t2<- c()
strata_y_fu_t1 <- c()
strata_y_fu_t2<- c()
strata_y_comp_n_events_t1 <- c()
strata_y_comp_n_events_t2 <- c()
strata_y_comp_fu_t1<- c()
strata_y_comp_fu_t2 <- c()

for(i in 1:length(strata_y)){
strata_y_n[i] <-length(strata_y[[i]]$i_date) 
strata_y_n_events_t1[i]<- sum(strata_y[[i]]$censor_all_str_t1)
strata_y_n_events_t2[i]<- sum(strata_y[[i]]$censor_all_str_t2)
strata_y_fu_t1[i] <- sum(strata_y[[i]]$survival_stroke_t1)
strata_y_fu_t2[i]<- sum(strata_y[[i]]$survival_stroke_t2)

strata_y_comp_n_events_t1[i]<- sum(strata_y[[i]]$comp_event_n_t1)
strata_y_comp_n_events_t2[i]<- sum(strata_y[[i]]$comp_event_n_t2)
strata_y_comp_fu_t1[i] <- sum(strata_y[[i]]$time_to_comp_event_t1)
strata_y_comp_fu_t2[i]<- sum(strata_y[[i]]$time_to_comp_event_t2)
}
strata_y_vars<- list(strata_y_n=strata_y_n, strata_y_n_events_t1=strata_y_n_events_t1, strata_y_n_events_t2=strata_y_n_events_t2, strata_y_fu_t1=strata_y_fu_t1, strata_y_fu_t2=strata_y_fu_t2, strata_y_comp_n_events_t1=strata_y_comp_n_events_t1, strata_y_comp_n_events_t2=strata_y_comp_n_events_t2, strata_y_comp_fu_t1=strata_y_comp_fu_t1, strata_y_comp_fu_t2=strata_y_comp_fu_t2)

#column names
strata <-c("Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")
strata_y <-c("Male <65 years", "Female <65 years", "Male 65-75 years", "Female 65-75 years","Male 75-85 years", "Female  75-85 years","Male 85 years and over", "Female 85 years and over")
```


```{r tables of n events for different strata, results='asis'}

strata_events <- do.call(cbind.data.frame, strata_y_vars)
setDT(strata_events)
strata_events <-strata_events[, strata_names:=strata_y][, c("strata_y_n", "strata_y_n_events_t1", "strata_y_n_events_t2","strata_y_comp_n_events_t1", "strata_y_comp_n_events_t2")]
strata_events<- cbind(strata_y, strata_events)
colnames(strata_events)<- c("Stratum", "N in strata", "Recurrent stroke by t1","Recurrent stroke by t2","Competing events by t1","Competing events by t2" )

ft <- flextable(strata_events)
ft<- width(ft,width=dim_pretty(ft)$widths)

flextable_to_rmd(ft)

cat("\n\n")

```

#### Event rates in LATCH over whole study time

```{r simple event rates}
## Set number of samples with nrep for use throughout. 
nrep<-1000

#event rate at t1 and t2
rate_est <- strata_y_vars$strata_y_n_events_t2/ strata_y_vars$strata_y_fu_t2*365.35*100
log_rate_est <- log(strata_y_vars$strata_y_n_events_t2/ strata_y_vars$strata_y_fu_t2)

# Calculate SE of the log-rate for untreated
log_rate_se <- 1 / sqrt(strata_y_vars$strata_y_n_events_t2)

## obtain and plot samples from rate in untreated
untreated_rates<- as.data.frame(cbind(strata_y_vars$strata_y_n, rate_est,log_rate_est, log_rate_se))
setDT(untreated_rates)
log_lambda_u<- untreated_rates[,rnorm(nrep, mean=log_rate_est, sd=log_rate_se), by=.I]

#rename I column to the strata names

log_lambda_u$I <- factor(log_lambda_u$I, levels=seq(1, length(strata_y)), labels=strata_y)

#calculate confidence intervals for visualisation
untreated_rates[, lower_ci:= exp(log_rate_est-1.96*log_rate_se)*365.35*100][, upper_ci:= exp(log_rate_est+1.96*log_rate_se)*365.35*100]
saveRDS(untreated_rates, "RDS files/LATCH untreated rates, strata above and below 65 years.rds")

#round all the numbers
untreated_rates_round<- round(untreated_rates,2)
#make the rate and confidence intervals look nice in one column
untreated_rates_round[, rate_cis:= paste(rate_est, " (", lower_ci, "-", upper_ci, ")", sep="")][, c("log_rate_est","rate_est","lower_ci","upper_ci","log_est","log_rate_se" ):= NULL][,Stratum:= strata_y ]
setnames(untreated_rates_round, c("V1", "rate_cis"), c("Recurrent stroke (n)","Rate of recurrent stroke per 100py (95%CI)"))

## competing risk rates
#event rate at t2 for each stratum, single value
rate_comp_est <- strata_y_vars$strata_y_comp_n_events_t2/ strata_y_vars$strata_y_comp_fu_t2*365.35*100
log_rate_comp_est <- log(strata_y_vars$strata_y_comp_n_events_t2/ strata_y_vars$strata_y_comp_fu_t2)

# Calculate SE of the log-rate for untreated, save
log_rate_comp_se <- 1 / sqrt(strata_y_vars$strata_y_comp_n_events_t2)

## obtain and plot samples from rate in untreated. This does nrep samples for each row (the .I tells it to do for each row). The output is nrep rows for each stratum, so rows 1:nrep are stratum 1, (nrep+1:2*nrep) are stratum 2, etc. This can be confirmed by checking the mean of each chunk of rows. 
untreated_rates_comp<- as.data.frame(cbind(strata_y_vars$strata_y_n, rate_comp_est,log_rate_comp_est, log_rate_comp_se))
setDT(untreated_rates_comp)
log_lambda_u_comp<- untreated_rates_comp[,rnorm(nrep, mean=log_rate_comp_est, sd=log_rate_comp_se), by=.I]

#rename I column to the strata names
strata <-c("Male <65 years", "Female <65 years", "Male 65 years and over", "Female 65 years and over")
log_lambda_u_comp$I <- factor(log_lambda_u_comp$I, levels=seq(1, length(strata_y)), labels=strata_y)

#calculate confidence intervals for visualisation
untreated_rates_comp[, lower_ci:= exp(log_rate_comp_est-1.96*log_rate_comp_se)*365.35*100][, upper_ci:= exp(log_rate_comp_est+1.96*log_rate_comp_se)*365.35*100]
saveRDS(untreated_rates_comp, "RDS files/LATCH untreated rates_comp, strata above and below 65 years.rds")

#round all the numbers
untreated_rates_comp_round<- round(untreated_rates_comp,2)
#make the rate and confidence intervals look nice in one column
untreated_rates_comp_round[, rate_cis:= paste(rate_comp_est, " (", lower_ci, "-", upper_ci, ")", sep="")][, c("rate_comp_est","lower_ci","upper_ci","log_rate_comp_est","log_rate_comp_se" ):= NULL][,Stratum:= strata_y ]
setnames(untreated_rates_comp_round, c("V1", "rate_cis"), c("Non-stroke death (n)", "Rate of non-stroke death per 100py (95%CI)"))

```


```{r table of rates}
#print table of rates

untreated_rates_all<- merge(untreated_rates_round, untreated_rates_comp_round, by="Stratum")
untreated_rates_all<- as.data.frame(untreated_rates_all)


ft_rates<- flextable(untreated_rates_all)

ft_rates2 <- width(ft_rates,width=dim_pretty(ft_rates)$widths)
ft_rates2

```


<!-- #### Plot of log event rates for each stratum -->
<!-- ```{r plot of event rates log,fig.width=8,fig.height=4} -->
<!-- #plot to confirm normality -->
<!-- ggplot(data=log_lambda_u)+ -->
<!--   geom_histogram(aes(x=V1), fill="#FCCDE5")+ -->
<!--   theme_bw()+ -->
<!--   facet_wrap(vars(I), nrow=2, ncol=4)+ -->
<!--   labs(title="Simulated log rates for strata 1-8")+  -->
<!--   xlab("") -->
<!-- ``` -->


### Hazard ratios for each stratum

These hazard ratios are extracted from TRIDENT.

```{r  hazard ratios trident}
trident_hrs_4_strata <- read_excel("~/Documents/REA3 clinical fellow job/Treatment effect in trials/TRIDENT information/trident_hrs_4_strata.xlsx")
setDT(trident_hrs_4_strata)
hrs_t1 <- trident_hrs_4_strata[, c("rec_stroke_hr_t1", "rec_stroke_l_ci_t1", "rec_stroke_u_ci_t1")]
hrs_t2 <- trident_hrs_4_strata[, c("rec_stroke_hr_t2", "rec_stroke_l_ci_t2", "rec_stroke_u_ci_t2")]
colnames(hrs_t1) <- c("HR", "l_ci", "u_ci")
colnames(hrs_t2) <- c("HR", "l_ci", "u_ci")

### create extra rows for over 65 to go with 65-75, 75-85, over 85
hrs_t1_over65 <-hrs_t1[3:4,] 
hrs_t1<- rbind(hrs_t1,hrs_t1_over65 ,hrs_t1_over65)
setDT(hrs_t1)

hrs_t2_over65 <-hrs_t2[3:4,] 
hrs_t2<- rbind(hrs_t2,hrs_t2_over65 ,hrs_t2_over65)
setDT(hrs_t2)

```


```{r hr for t1}
hrs_t1_for_tab <-round(hrs_t1,2)

hrs_t1_for_tab[, Stratum:= strata_y][, "HR (95%CI) T1":= paste(HR, " (", l_ci, "-", u_ci, ")", sep="")][,c("HR","l_ci","u_ci"):=NULL ]

hrs_t2_for_tab <-round(hrs_t2,2)
hrs_t2_for_tab[, Stratum:= strata_y][, "HR (95%CI) T2":= paste(HR, " (", l_ci, "-", u_ci, ")", sep="")][,c("HR","l_ci","u_ci"):=NULL ]

hrs_both_times_tab<- merge(hrs_t1_for_tab, hrs_t2_for_tab, by="Stratum")
ft_hr_t1<- flextable(hrs_both_times_tab)
ft_hr_t1<- set_header_labels(ft_hr_t1, HR="Hazard ratio", l_ci="Lower CI", u_ci="Upper CI")
ft_hr_t1_a <- width(ft_hr_t1,width=dim_pretty(ft_hr_t1)$widths)
ft_hr_t1_a



```

<!-- #### Histogram of log hazard ratios for strata 1-8, $\tau_1$ -->

```{r 1000 hrs, warnings=F,fig.width=8,fig.height=4}
#obtain log of hrs and CI at t1 and t2. Assume log normally distributed, and find log standard error
setDT(hrs_t1)
setDT(hrs_t2)
hrs_t1[, log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]
hrs_t2[, log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]

#simulate log hr for nrep samples.
#create simulated hr for all strata
log_hrs_t1_per_strata <- hrs_t1[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]

#rename I column to the strata names
log_hrs_t1_per_strata$I <- factor(log_hrs_t1_per_strata$I, levels=seq(1, length(strata_y)), labels=strata_y)

#plot to confirm normality
# ggplot(data=log_hrs_t1_per_strata)+
#   geom_histogram(aes(x=V1), fill="#FCCDE5")+
#   theme_bw()+
#   facet_wrap(vars(I), nrow=2, ncol=4)+
#   labs(title="Simulated log hazard ratios for strata, T1")+
#   xlab("Log of hazard ratio")
```

<!-- #### Histogram of simulated hazard ratios for strata 1-8, $\tau_2$ -->

```{r hist hr t1 exponentiated,fig.width=8,fig.height=4}
# ggplot(data=log_hrs_t1_per_strata)+
#   geom_histogram(aes(x=exp(V1)), fill="#FCCDE5")+
#   theme_bw()+
#   facet_wrap(vars(I), nrow=2, ncol=4)+
#   labs(title="Simulated hazard ratios for strata, T1")+
#   xlab("Hazard ratio")
```

<!-- #### Histogram of simulated log hazard ratios for strata, $\tau_2$ -->

```{r t2 hr hists,fig.width=8,fig.height=4}
 #repeat for t2 hazard ratios 
#create simulated hr for all strata
log_hrs_t2_per_strata <- hrs_t2[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]

 #rename I column to the strata names
log_hrs_t2_per_strata$I <- factor(log_hrs_t2_per_strata$I, levels=seq(1, length(strata_y)), labels=strata_y)

# #plot to confirm normality
# ggplot(data=log_hrs_t2_per_strata)+
#   geom_histogram(aes(x=V1), fill="#80B1D3")+
#   theme_bw()+
#   facet_wrap(vars(I), nrow=2, ncol=4)+
#   labs(title="Simulated log hazard ratios for strata, T2")+
#   xlab("Log of hazard ratio")
```

<!-- #### Histogram of simulated hazard ratios for strata 1-8, $\tau_2$ -->

```{r t2 hr exponentiated,fig.width=8,fig.height=4}
# ggplot(data=log_hrs_t2_per_strata)+
  # geom_histogram(aes(x=exp(V1)), fill="#80B1D3")+
  # theme_bw()+
  # facet_wrap(vars(I), nrow=2, ncol=4)+
  # labs(title="Simulated hazard ratios for strata, T2")+
  # xlab("Hazard ratio")
```

### Calibrated rates and CIs for each stratum {.tabset}

The histograms show the spread of risk ratio for 1000 simulated samples, at t1 and t2. I have marked on the mean for each stratum.

```{r risks treated and untreated 2}
#set t1 and t2 in days
t1<-548
t2<- 4.5*365.25

#lambda from parametric model above
#lambda_u <- exp(z_lambda) 
log_lambda_u_sim<- log_lambda_u$V1

#for rate in treated, log lambda_t = log lambda_u + log hr
log_lambda_t1_sim <- log_lambda_u_sim+log_hrs_t1_per_strata$V1
log_lambda_t2_sim <- log_lambda_u_sim+log_hrs_t2_per_strata$V1

#exponentiate to get rates
lambda_u_sim <- exp(log_lambda_u_sim)
lambda_t1_sim<- exp(log_lambda_t1_sim)
lambda_t2_sim<- exp(log_lambda_t2_sim)

#transform to risks
risk_u1 <- 1-exp(-lambda_u_sim*t1)
risk_t1 <- 1-exp(-lambda_t1_sim*t1)
risk_u2 <- 1-exp(-lambda_u_sim*(t1+t2))
risk_t2 <- 1-exp(-(lambda_t1_sim*t1+lambda_t2_sim*t2))
risk_ratio_t1 <- risk_t1/risk_u1
risk_ratio_t2 <- risk_t2/risk_u2

risk_ratio_for_plot_t1 <- as.data.frame(cbind(rep(seq(1, length(strata_y)), each=nrep), risk_ratio_t1))
risk_ratio_for_plot_t2 <- as.data.frame(cbind(rep(seq(1, length(strata_y)), each=nrep), risk_ratio_t2))

rr_means_t1<- risk_ratio_for_plot_t1%>%
  group_by(V1)%>%
  summarize(means_t1=round(mean(risk_ratio_t1),3))
  
rr_means_t2<- risk_ratio_for_plot_t2%>%
  group_by(V1)%>%
  summarize(means_t2=round(mean(risk_ratio_t2),3))

```

#### Risk ratios for strata, $t_1$

```{r risk plot t1,fig.width=8,fig.height=4}
ggplot(data=risk_ratio_for_plot_t1)+
  geom_histogram(aes(x=risk_ratio_t1),breaks = seq(0, 4, by = 0.1), fill="#FCCDE5")+
  geom_text(data=rr_means_t1, aes(x=means_t1+1, y=100, label= paste("Mean=",means_t1)), size=3)+
  theme_bw()+
  geom_vline(aes(xintercept = means_t1, colour=V1), linewidth=0.5, linetype="dashed", data=rr_means_t1)+
  facet_wrap(vars(V1), nrow=2, ncol=4)+
  labs(title=quote("Simulated risk ratios for strata, "* t[1]* "."))+
  xlab("Risk ratio")+
  ylab("Number of samples")+
  xlim(c(0,4))+
  ylim(c(0,150))

```


#### Risk ratios for strata 1-8, $t_2$

```{r risk plot t2,fig.width=8,fig.height=4}
ggplot(data=risk_ratio_for_plot_t2)+
  geom_histogram(aes(x=risk_ratio_t2), fill="#80B1D3")+
  geom_text(data=rr_means_t2, aes(x=means_t2+1, y=100, label= paste("Mean=",means_t2)), size=3)+
  theme_bw()+
  geom_vline(aes(xintercept = means_t2, colour=V1), linewidth=0.5, linetype="dashed", data=rr_means_t2)+
  facet_wrap(vars(V1), nrow=2, ncol=4)+
  labs(title= quote("Simulated risk ratios for strata 1-8, "* t[2]* "."))+
  xlab("Risk ratio")

```

### Pooled risk ratio

```{r pooling hrs}
#matrix of risk ratios at t1, one column per stratum

rr_t1<-matrix(risk_ratio_t1, ncol=length(strata_y))
rr_t2<-matrix(risk_ratio_t2, ncol=length(strata_y))

#matrix of ns in latch, one column per stratum
strata_n_df<- as.data.frame(t(strata_y_vars$strata_y_n))
ns <- as.matrix(strata_n_df %>% slice(rep(1, each = nrep)))
#create a vector of totals
n_tot<- rowSums(ns)

## calculate n x rr_mtrx (note that this is NOT matrix multiplication)
#t1
rn_t1 <- rowSums(rr_t1*ns)
rr_tot_t1 <- rn_t1/n_tot
cis_t1<- round(quantile(rr_tot_t1, probs = c(0.025, 0.975)),3)
est_t1<- round(mean(rr_tot_t1),3)

#t2
rn_t2 <- rowSums(rr_t2*ns)
rr_tot_t2 <- rn_t2/n_tot
cis_t2<- round(quantile(rr_tot_t2, probs = c(0.025, 0.975)),3)
est_t2<- round(mean(rr_tot_t2),3)
```

#### Calibrated risk ratio at the end of the trial, for the whole trial

```{r t2 pooled results numerical, comment=NA}

cat(paste("  The estimated calibrated risk ratio is ", est_t2, " (95% confidence interval ", cis_t2[1], "-", cis_t2[2], ").", sep=""))

```
\n
\n
```{r t2 pooled results,results="asis",fig.width=4,fig.height=3}

ggplot(data=as.data.frame(rr_tot_t2))+
  geom_histogram(aes(x=rr_tot_t2), fill="#80B1D3")+
  theme_bw()+
  labs(title=quote("Simulated risk ratio for the whole trial at "* t[2]* ". "))+ 
  geom_vline(aes(xintercept = est_t2), linewidth=0.5, linetype="dashed")+
  xlab("Risk ratio")+  
  ylab("n samples")+
  geom_text(data=rr_means_t1, aes(x=est_t2+0.16, y=75, label= paste("RR=",est_t1," (95%CI ", cis_t2[1], "-", cis_t2[2], ")", sep="")), size=3)



```

### Number needed to treat
  
```{r absolute risks, comment=NA}
#without simulation
event_rate<- sum(strata_vars$strata_n_events_t2)/sum(strata_vars$strata_fu_t2)*365.25
arr <- event_rate- event_rate*rr_tot_t2
nnt <- 1/arr
cis_nnt<- quantile(nnt, probs = c(0.025, 0.975))
cis_arr<- round(quantile(arr, probs = c(0.025, 0.975)),4)

#with simulation
# event_rate_sim<- lambda_u_sim*365.25
# arr_sim <- (event_rate_sim- event_rate_sim*rr_tot_t2)
# nnt_sim<- 1/arr_sim
# cis_nnt_sim<- quantile(nnt_sim, probs = c(0.025, 0.975))
# cis_arr_sim<- round(quantile(arr_sim, probs = c(0.025, 0.975)),4)

cat(paste(" The number needed to treat is ", round(mean(nnt),2)," (95% CI ", round(cis_nnt[1],2), " - ", round(cis_nnt[2],2),").", sep=""))

```

### Competing event analysis: assuming no difference in non-stroke death between treatment and placebo arm

```{r loop through strata for arr comp 1, results='asis'}

#We want arr for each stratum. The lambdas are in a single vector, length of the number of strata time the number of reps.

t1_times <- seq(0, 548, 1)
t2_times <- seq(549, 365.25*6, 1)
all_times <- c(t1_times, t2_times)

pl_2 <- list()
pl_2_cis <- list()
tx_2 <- list()
tx_2_cis <- list()

arr_2 <- list()
rr_2 <- list()
or_2<- list()

for(i in 1:length(strata_y)){
# Placebo arm
pl_2[[i]] <- MakeCumulative2(arm="u",
                       rate.main.t1 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)],
                       rate.main.t2 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)],
                       rate.compete.t1 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]),
                       rate.compete.t2 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]),
                       t1_times,
                       t2_times)
pl_2_cis[[i]]<- get_CIs(pl_2[[i]], c(t1_times, t2_times))

# Treatment arm
tx_2[[i]] <- MakeCumulative2 (arm = "rx",
                        rate.main.t1 = lambda_t1_sim[(nrep*(i-1)+1):(i*nrep)],
                        rate.main.t2 = lambda_t2_sim[(nrep*(i-1)+1):(i*nrep)],
                        rate.compete.t1 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]),
                        rate.compete.t2 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]),
                        t1_times,
                        t2_times)
tx_2_cis[[i]]<- get_CIs(tx_2[[i]],c(t1_times, t2_times))
## Calculate comparisons
arr_2[[i]] = pl_2[[i]] - tx_2[[i]]
rr_2[[i]] = tx_2[[i]]/pl_2[[i]]
or_2[[i]] = (tx_2[[i]]/(1-tx_2[[i]])) / (pl_2[[i]]/(1-pl_2[[i]]))

}

```

#### Plots of outcomes {.tabset}

These plots show the cumulative incidence of a) recurrent stroke, b) death due to none-stroke cause and c) recurrent stroke, accounting for competing risk of death.

##### Recurrent stroke outcome 

```{r plots of outcomes comp 1, fig.width=12, fig.height=6, results='asis'}
plots_rec_str<- list()
arms_x1<- list()
for(i in 1:length(strata_y)){
  
  arms_x1[[i]] <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_main, l_ci = pl_2_cis[[i]]$l_ci_main, u_ci=pl_2_cis[[i]]$u_ci_main),
                      tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_main, l_ci = tx_2_cis[[i]]$l_ci_main, u_ci=tx_2_cis[[i]]$u_ci_main))
  
  ## Create plots
  setDT(arms_x1[[i]])
  plots_rec_str[[i]] <- ggplot(arms_x1[[i]], aes(x = obs.times, y = n1_cum_per, colour = arm)) +
      geom_ribbon(aes(ymin = l_ci, ymax = u_ci, colour=NULL, fill=arm), alpha=0.2) +
      geom_line()+
      #geom_point()+
      scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
                     labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
      scale_y_continuous ("Cumulative incidence", limits=c(0,0.5)) +
      theme_bw()+
      theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+
      ggtitle(paste(strata_y[i]))+
      guides(color="none")+
      labs(fill="Arm")
}

(plots_rec_str[[1]]|plots_rec_str[[2]]|plots_rec_str[[3]]|plots_rec_str[[4]])/(plots_rec_str[[5]]|plots_rec_str[[6]]|plots_rec_str[[7]]|plots_rec_str[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke for each stratum")

```

##### Competing risk outcome 

```{r competing outcome plots comp 1,fig.width=12, fig.height=6, results='hide'}
plots_comp = list()
for(i in 1:length(strata_y)){

arms_x2 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_comp, l_ci = pl_2_cis[[i]]$l_ci_comp, u_ci=pl_2_cis[[i]]$u_ci_comp),
               tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_comp, l_ci = tx_2_cis[[i]]$l_ci_comp, u_ci=tx_2_cis[[i]]$u_ci_comp))
## set upper limit on upper ci of 1 so it plots properly

arms_x2<- arms_x2%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci))
## Create plots
plots_comp[[i]] <- ggplot(arms_x2, aes(x = obs.times, y = n1_cum_per, colour = arm)) +
    geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) +
    geom_line()+
    scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
                   labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
    scale_y_continuous ("Cumulative incidence", limits=c(0,1)) +
    theme_bw()+
    theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+
    ggtitle(paste(strata_y[i]))+
    guides(color="none")+
    labs(fill="Arm")
}

#(plots_comp[[1]]|plots_comp[[2]]|plots_comp[[3]]|plots_comp[[4]])/(plots_comp[[5]]|plots_comp[[6]]|plots_comp[[7]]|plots_comp[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of competing risk for each stratum")

```

##### Recurrent stroke, accounting for competing risk of non-stroke death 

```{r both outcome plots comp 1,fig.width=12, fig.height=6,  results='hide'}

plot_both= list()
for(i in 1:length(strata_y)){
  
  arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both),
                 tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both))
  
  
  ## set upper limit on upper ci of 1 so it plots properly
  
  arms_x3<- arms_x3%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci))
  ## Create plots
  plot_both[[i]] <- ggplot(arms_x3, aes(x = obs.times, y = n1_cum_per, colour = arm)) +
  geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) +
  geom_line()+
  scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
                 labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
  scale_y_continuous ("Cumulative incidence", limits=c(0,1)) +
  theme_bw()+
  theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+
  ggtitle(paste(strata_y[i]))+
  guides(color="none")+
  labs(fill="Arm")

}

(plot_both[[1]]|plot_both[[2]]|plot_both[[3]]|plot_both[[4]])/(plot_both[[5]]|plot_both[[6]]|plot_both[[7]]|plot_both[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke, accounting for competing risk of death")

```

#### Absolute risk reduction {.tabset}

These plots and tables show the absolute risk reduction for each stratum and overall, accounting for competing risk of death.

```{r format ARR comp 1, results='asis'}
#create dataframes and tables of arr and event rates in each arm, for each stratum
arr_cis<- list() 
for(i in 1:length(strata_y)){
  
  arr_cis[[i]] <- as.data.frame(get_CIs(arr_2[[i]], all_times))
  setDT(arr_cis[[i]])
  arr_cis[[i]][, time:= all_times]
  arr_cis_select<- arr_cis[[i]][time %in% times_for_tables][,c("time","mean_both", "l_ci_both", "u_ci_both") ]
  arr_cis_select[, mean_both:= round(mean_both*100,1)][, u_ci_both:=round(u_ci_both*100,1)][, l_ci_both:=round(l_ci_both*100,1)]
  
  colnames(arr_cis_select) <- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)")
  
  arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "pl", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both),
                 tibble(obs.times = c(t1_times, t2_times), arm = "tx", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both))
  
  setDT(arms_x3)
  arms3_select <- arms_x3[obs.times %in% times_for_tables]
  abs_rates <- dcast(arms3_select, obs.times ~ arm, value.var="n1_cum_per")
  abs_rates[,pl:= round(100*pl,1)][, tx:= round(100*tx,1)]
  colnames(abs_rates) <- c("Time (days)", "Events in untreated (%)", "Events in treated (%)")
  
  arr_and_individ <- merge(abs_rates, arr_cis_select, by="Time (days)")
  ft<- flextable(arr_and_individ)
  ft<- width(ft,width=dim_pretty(ft)$widths)
  
  cat(paste0('##### ', strata_y[i], '\n\n'))
  
  flextable_to_rmd(ft)
  
  cat("\n\n")
}
```

##### Overall

```{r arr in one table comp 1, results='asis'}

#calculate ARR for whole cohort
#matrix of risk ratios at t1, one column per stratum, nrep rows for each day. Need to do this once for each day.
arr_per_strata <- unlist(lapply(arr_2, function(x) x[,V3, drop=FALSE]))

arr_flat<-matrix(arr_per_strata, ncol=length(strata_y))

#matrix of ns in latch, one column per stratum
strata_n_df<- as.data.frame(t(strata_y_vars$strata_y_n))
ns <- as.matrix(strata_n_df %>% slice(rep(1, each = nrep*length(all_times))))
#create a vector of totals
n_tot<- rowSums(ns)

arr_n <- rowSums(arr_flat*ns)
arr_n_tot <- as.data.frame(arr_n/n_tot)

setDT(arr_n_tot)
arr_n_tot[, index:=rep(all_times, each=nrep)]

#Group by index then work out quantiles. Use data.table
arr_n_tot[,mean_both:= (mean(`arr_n/n_tot`,na.rm=T)*100), by=index][,l_ci_both:=(quantile(`arr_n/n_tot`, probs=0.025, na.rm=T)*100),by=index][, u_ci_both:=(quantile(`arr_n/n_tot`, probs=0.975,na.rm=T)*100), by=index]

#select every nrep row
arr_n_tot_new <- arr_n_tot[seq(1, nrow(arr_n_tot),nrep), ]

#a note on the formatting of the risks outputs. For ease of calculating the confidence intervals (to keep compatible with the function get_CIs), the data frames include unnecessary columns. We select the relevant ones now.

arr_rel_cols<- list()
for(i in 1: length(strata_y)){
setDT(arr_cis[[i]])
arr_rel_cols[[i]]<-  arr_cis[[i]][,c("time", "mean_both", "l_ci_both", "u_ci_both")][, strata:=strata_y[i]][, mean_both:= 100*mean_both][, l_ci_both:= 100*l_ci_both] [,u_ci_both:= 100*u_ci_both]
}

arr_all <- do.call(rbind.data.frame, arr_rel_cols)
#we want to add in rows for the combined weighted total. First harmonise column names

arr_comb_1 <- arr_n_tot_new[, c("index","mean_both","l_ci_both","u_ci_both")][,strata:="All"]
setnames(arr_comb_1, "index", "time")

arr_all_comb<- rbind(arr_all, arr_comb_1)

### table at various time points
arr_select <- arr_n_tot_new[index %in% times_for_tables][, `arr_n/n_tot`:=NULL]
colnames(arr_select)<- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)")

ft_comb<- flextable(round(arr_select,1))
ft_comb<- width(ft_comb,width=dim_pretty(ft_comb)$widths)

ft_comb
```

#### Plots {.tabset}



##### Plot of ARR for each strata

```{r plot overall comp 1, results='asis', fig.height=4, fig.width=9}

#all on one plot
p_1 <- ggplot(data=arr_all)+
geom_smooth( aes(x =time, y = mean_both, colour=strata), linewidth=0.5, linetype="dashed")+
geom_smooth( data=arr_comb_1,aes(x =time, y = mean_both, colour=strata), linewidth=1)+
geom_ribbon(data=arr_comb_1, aes(ymin = l_ci_both, ymax = u_ci_both, x=time),fill= "#377EB8",alpha=0.1)+ 
scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
             labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
scale_y_continuous ("Absolute risk reduction (%)") +
scale_color_discrete(palette=c("#377EB8","#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+
theme_bw()+
ggtitle(paste("Absolute risk reduction of recurrent stroke, HR for competing risk of death = 1"))+
labs(color="Stratum")
p_1
ggsave("ARR plots with varying competing risks/Competing risk ratio 1, combined.png")

```

##### Plot of ARR for each stratum separately

Absolute risk reduction for the whole LATCH population and each age/sex stratum, taking into account competing risk of death.
```{r plot sep comp 1, fig.height=4, fig.width=12}
#strata separately
ggplot(data=arr_all)+
geom_smooth( aes(x =time, y = mean_both,color=strata), linewidth=0.5, linetype="dashed")+
geom_ribbon( aes(ymin = l_ci_both, ymax = u_ci_both, x=time,fill=strata),alpha=0.1)+ 
facet_wrap(facets="strata", ncol=4, nrow=2)+
scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
             labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
scale_y_continuous ("Absolute risk reduction (%)") +
scale_color_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+
scale_fill_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+
theme_bw()+
ggtitle(paste("Absolute risk reduction of recurrent stroke, HR for competing risk of death = 1"))+
labs(color="Stratum")


ggsave("ARR plots with varying competing risks/Competing risk ratio 1, separate.png")

```

### Competing event analysis: assuming difference in non-stroke death between treatment and placebo arm using hazard ratios from TRIDENT data.

#### Hazard ratios from TRIDENT data

```{r table of TRIDENT HR}
#get the real competing event rates from each arm.
HR_comp_t1 <- trident_hrs_4_strata[, c("comp_hr_t1","comp_lci_t1","comp_uci_t1" )]
HR_comp_t2 <- trident_hrs_4_strata[, c("comp_hr_t2","comp_lci_t2","comp_uci_t2" )]
colnames(HR_comp_t1)<- c("HR", "l_ci", "u_ci")
colnames(HR_comp_t2)<- c("HR", "l_ci", "u_ci")

### create extra rows for over 65 to go with 65-75, 75-85, over 85
HR_comp_t1_over65 <-HR_comp_t1[3:4,] 
HR_comp_t1<- rbind(HR_comp_t1,HR_comp_t1_over65 ,HR_comp_t1_over65)
setDT(HR_comp_t1)

HR_comp_t2_over65 <-HR_comp_t2[3:4,] 
HR_comp_t2<- rbind(HR_comp_t2,HR_comp_t2_over65 ,HR_comp_t2_over65)
setDT(HR_comp_t2)

### make version to print
hrs_comp_t1_for_tab <-round(hrs_t1,2)

hrs_comp_t1_for_tab<-hrs_comp_t1_for_tab[, Stratum:= strata_y][, "HR (95%CI) T1":= paste(HR, " (", l_ci, "-", u_ci, ")", sep="")][,c("HR","l_ci","u_ci", "log_se", "log_hr", "se"):=NULL ]

hrs_comp_t2_for_tab <-round(hrs_t2,2)
hrs_comp_t2_for_tab <- hrs_comp_t2_for_tab[, Stratum:= strata_y][, "HR (95%CI) T2":= paste(HR, " (", l_ci, "-", u_ci, ")", sep="")][,c("HR","l_ci","u_ci", "log_se", "log_hr", "se"):=NULL ]

hrs_comp_both_times_tab<- merge(hrs_comp_t1_for_tab, hrs_comp_t2_for_tab, by="Stratum")

hrs_comp_both_times_tab$Stratum<- factor(hrs_comp_both_times_tab$Stratum, levels=strata_y, ordered=T)
hrs_comp_both_times_tab<- hrs_comp_both_times_tab%>%arrange(hrs_comp_both_times_tab, hrs_comp_both_times_tab$Stratum)

ft_hr_t1<- flextable(hrs_comp_both_times_tab)
#ft_hr_t1<- set_header_labels(ft_hr_t1, HR="Hazard ratio", l_ci="Lower CI", u_ci="Upper CI")
ft_hr_t1_a <- width(ft_hr_t1,width=dim_pretty(ft_hr_t1)$widths)
ft_hr_t1_a


```

#### Event rates for competing risk using hazard ratios from TRIDENT

```{r real competing event rate in rx arm}
#get the real competing event rates from each arm.
HR_comp_t1 <- trident_hrs_4_strata[, c("comp_hr_t1","comp_lci_t1","comp_uci_t1" )]
HR_comp_t2 <- trident_hrs_4_strata[, c("comp_hr_t2","comp_lci_t2","comp_uci_t2" )]
colnames(HR_comp_t1)<- c("HR", "l_ci", "u_ci")
colnames(HR_comp_t2)<- c("HR", "l_ci", "u_ci")

### create extra rows for over 65 to go with 65-75, 75-85, over 85
HR_comp_t1_over65 <-HR_comp_t1[3:4,] 
HR_comp_t1<- rbind(HR_comp_t1,HR_comp_t1_over65 ,HR_comp_t1_over65)
setDT(HR_comp_t1)

HR_comp_t2_over65 <-HR_comp_t2[3:4,] 
HR_comp_t2<- rbind(HR_comp_t2,HR_comp_t2_over65 ,HR_comp_t2_over65)
setDT(HR_comp_t2)

# calculate the log hr, and log se from normal distribution formulae
HR_comp_t1[,log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]
HR_comp_t2[,log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)]

# simulate nrep samples. .I means, do each row nrep times. The output is nrep rows for each stratum, so rows 1:nrep are stratum 1, (nrep+1:2*nrep) are stratum 2, etc. This can be confirmed by checking the mean of each chunk of rows. 
log_hr_comp_t1<- HR_comp_t1[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]
log_hr_comp_t2<- HR_comp_t2[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I]

#for each time period, generate a rate of competing risk in the treated.
log_lambda_t_comp_t1 <- copy(log_lambda_u_comp)
log_lambda_t_comp_t2 <- copy(log_lambda_u_comp)
log_lambda_t_comp_t1[, log_rate_comp_t_est_t1:=V1+log_hr_comp_t1$V1 ][,V1:=NULL]
log_lambda_t_comp_t2[, log_rate_comp_t_est_t2:=V1+log_hr_comp_t2$V1 ][,V1:=NULL]
#harmonise names with untreated
setnames(log_lambda_t_comp_t1, "log_rate_comp_t_est_t1", "V1")
setnames(log_lambda_t_comp_t2, "log_rate_comp_t_est_t2", "V1")

#calculate confidence intervals for visualisation
conf_int_treated_t1 <- log_lambda_t_comp_t1[,quantile(V1, c(0.025,0.5,0.975)), by=I]
conf_int_treated_t1[, cat:= rep(c("Lower CI", "Mean", "Upper CI"), length(strata_y))]

conf_int_treated_t2 <- log_lambda_t_comp_t2[,quantile(V1, c(0.025,0.5,0.975)), by=I]
conf_int_treated_t2[, cat:= rep(c("Lower CI", "Mean", "Upper CI"), length(strata_y))]

#widen table
treated_comp_t1 <-dcast(conf_int_treated_t1, I~cat, value.var = "V1")
treated_comp_t1_x<- treated_comp_t1[, "Rate per 100 person/years":=paste(round(exp(Mean)*100*365.25,2), " (",round(exp(`Lower CI`)*100*365.25,2), " - ",round(exp(`Upper CI`)*100*365.25,2),")", sep="") ][, "Lower CI":=NULL][, "Upper CI":=NULL]
colnames(treated_comp_t1_x) <- c( "Stratum",  "log(rate/day)","Rate of non-stroke death per 100py (95%CI)")

treated_comp_t2 <-dcast(conf_int_treated_t2, I~cat, value.var = "V1")
treated_comp_t2_x<- treated_comp_t2[, "Rate per 100 person/years":=paste(round(exp(Mean)*100*365.25,2), " (",round(exp(`Lower CI`)*100*365.25,2), " - ",round(exp(`Upper CI`)*100*365.25,2),")", sep="") ][, "Lower CI":=NULL][, "Upper CI":=NULL]
colnames(treated_comp_t2_x) <- c( "Stratum",  "log(rate/day)","Rate of non-stroke death per 100py (95%CI)")

# table of rates combined
setDT(untreated_rates_comp_round)
unt_for_merge <- untreated_rates_comp_round[, Stratum:=strata_y][, c("Stratum", "Rate of non-stroke death per 100py (95%CI)")]
tr_comp_t1_for_m<- treated_comp_t1_x[, c("Stratum", "Rate of non-stroke death per 100py (95%CI)")]
tr_comp_t2_for_m<- treated_comp_t2_x[, c("Stratum", "Rate of non-stroke death per 100py (95%CI)")]
#setnames(unt_for_merge, "rate_cis", "Rate of non-stroke death per 100py (95%CI)")
setnames(tr_comp_t1_for_m, "Rate of non-stroke death per 100py (95%CI)", "Events per 100py, treated t1")
setnames(tr_comp_t2_for_m, "Rate of non-stroke death per 100py (95%CI)", "Events per 100py, treated t2")
m_1<- merge(unt_for_merge, tr_comp_t1_for_m, by="Stratum")
rates_comp_comb<- merge(m_1, tr_comp_t2_for_m, by="Stratum")
rates_comp_comb$Stratum<- factor(rates_comp_comb$Stratum, levels=strata_y, ordered=T)
rates_comp_comb<- rates_comp_comb%>% arrange(Stratum)
rm(m_1, unt_for_merge, tr_comp_t1_for_m, tr_comp_t2_for_m)

ft_rates_comp<- flextable(rates_comp_comb)
ft_rates_comp <- width(ft_rates_comp,width=dim_pretty(ft_rates_comp)$widths)
ft_rates_comp

```


```{r loop through strata for arr, results='asis'}

#We want arr for each stratum. The lambdas are in a single vector, length of the number of strata time the number of reps.
pl_2 <- list()
pl_2_cis <- list()
tx_2 <- list()
tx_2_cis <- list()

arr_2 <- list()
rr_2 <- list()
or_2<- list()

for(i in 1:length(strata_y)){
  # Placebo arm
pl_2[[i]] <- MakeCumulative2(arm="u",
                       rate.main.t1 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)],
                       rate.main.t2 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)],
                       rate.compete.t1 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]),
                       rate.compete.t2 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]),
                       t1_times,
                       t2_times)
pl_2_cis[[i]]<- get_CIs(pl_2[[i]], c(t1_times, t2_times))

# Treatment arm
tx_2[[i]] <- MakeCumulative2 (arm = "rx",
                        rate.main.t1 = lambda_t1_sim[(nrep*(i-1)+1):(i*nrep)],
                        rate.main.t2 = lambda_t2_sim[(nrep*(i-1)+1):(i*nrep)],
                        rate.compete.t1 = exp(log_lambda_t_comp_t1[(nrep*(i-1)+1):(i*nrep),2]),
                        rate.compete.t2 = exp(log_lambda_t_comp_t2[(nrep*(i-1)+1):(i*nrep),2]),
                        t1_times,
                        t2_times)
tx_2_cis[[i]]<- get_CIs(tx_2[[i]],c(t1_times, t2_times))
## Calculate comparisons
arr_2[[i]] = pl_2[[i]] - tx_2[[i]]
rr_2[[i]] = tx_2[[i]]/pl_2[[i]]
or_2[[i]] = (tx_2[[i]]/(1-tx_2[[i]])) / (pl_2[[i]]/(1-pl_2[[i]]))
}

```

#### Plots of outcomes {.tabset}

These plots show the cumulative incidence of a) recurrent stroke, b) death due to none-stroke cause and c) recurrent stroke, accounting for competing risk of death.

##### Recurrent stroke outcome 

```{r plots of outcomes, fig.width=12, fig.height=6, results='asis'}
plots_rec_str<- list()
arms_x1<- list()
for(i in 1:length(strata_y)){
  
  arms_x1[[i]] <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_main, l_ci = pl_2_cis[[i]]$l_ci_main, u_ci=pl_2_cis[[i]]$u_ci_main),
                      tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_main, l_ci = tx_2_cis[[i]]$l_ci_main, u_ci=tx_2_cis[[i]]$u_ci_main))
  
  ## Create plots
  setDT(arms_x1[[i]])
  plots_rec_str[[i]] <- ggplot(arms_x1[[i]], aes(x = obs.times, y = n1_cum_per, colour = arm)) +
  geom_ribbon(aes(ymin = l_ci, ymax = u_ci, colour=NULL, fill=arm), alpha=0.2) +
  geom_line()+
  #geom_point()+
  scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
                 labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
  scale_y_continuous ("Cumulative incidence", limits=c(0,0.5)) +
  theme_bw()+
  theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+
  ggtitle(paste(strata_y[i]))+
  guides(color="none")+
  labs(fill="Arm")
}

(plots_rec_str[[1]]|plots_rec_str[[2]]|plots_rec_str[[3]]|plots_rec_str[[4]])/(plots_rec_str[[5]]|plots_rec_str[[6]]|plots_rec_str[[7]]|plots_rec_str[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke for each stratum")

```

##### Competing outcome (none-stroke death) 

```{r competing outcome plots,fig.width=12, fig.height=6, results='asis'}
plots_comp = list()
for(i in 1:length(strata_y)){
  
  arms_x2 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_comp, l_ci = pl_2_cis[[i]]$l_ci_comp, u_ci=pl_2_cis[[i]]$u_ci_comp),
                 tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_comp, l_ci = tx_2_cis[[i]]$l_ci_comp, u_ci=tx_2_cis[[i]]$u_ci_comp))
  ## set upper limit on upper ci of 1 so it plots properly
  
  arms_x2<- arms_x2%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci))
  ## Create plots
  plots_comp[[i]] <- ggplot(arms_x2, aes(x = obs.times, y = n1_cum_per, colour = arm)) +
  geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) +
  geom_line()+
  scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
                 labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
  scale_y_continuous ("Cumulative incidence", limits=c(0,1)) +
  theme_bw()+
  theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+
  ggtitle(paste(strata_y[i]))+
  guides(color="none")+
  labs(fill="Arm")


}

(plots_comp[[1]]|plots_comp[[2]]|plots_comp[[3]]|plots_comp[[4]])/(plots_comp[[5]]|plots_comp[[6]]|plots_comp[[7]]|plots_comp[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of competing risk for each stratum")

```

##### Recurrent stroke, accounting for competing risk of none-stroke death 

```{r both outcome plots,fig.width=12, fig.height=6,  results='asis'}

plot_both= list()
for(i in 1:length(strata_y)){
  
  arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both),
                 tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both))
  
  
  ## set upper limit on upper ci of 1 so it plots properly
  
  arms_x3<- arms_x3%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci))
  ## Create plots
  plot_both[[i]] <- ggplot(arms_x3, aes(x = obs.times, y = n1_cum_per, colour = arm)) +
  geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) +
  geom_line()+
  scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
                 labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
  scale_y_continuous ("Cumulative incidence", limits=c(0,1)) +
  theme_bw()+
  theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+
  ggtitle(paste(strata_y[i]))+
  guides(color="none")+
  labs(fill="Arm")

}

(plot_both[[1]]|plot_both[[2]]|plot_both[[3]]|plot_both[[4]])/(plot_both[[5]]|plot_both[[6]]|plot_both[[7]]|plot_both[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke, accounting for competing risk of death")

```

#### Absolute risk reduction {.tabset}

These plots and tables show the absolute risk reduction for each stratum and overall, accounting for competing risk of death.

```{r format ARR, results='asis'}
#create dataframes and tables of arr and event rates in each arm, for each stratum
arr_cis<- list() 


for(i in 1:length(strata_y)){
  
  arr_cis[[i]] <- as.data.frame(get_CIs(arr_2[[i]], all_times))
  setDT(arr_cis[[i]])
  arr_cis[[i]][, time:= all_times]
  arr_cis_select<- arr_cis[[i]][time %in% times_for_tables][,c("time","mean_both", "l_ci_both", "u_ci_both") ]
  arr_cis_select[, mean_both:= round(mean_both*100,1)][, u_ci_both:=round(u_ci_both*100,1)][, l_ci_both:=round(l_ci_both*100,1)]
  
  colnames(arr_cis_select) <- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)")
  
  arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "pl", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both),
                 tibble(obs.times = c(t1_times, t2_times), arm = "tx", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both))
  
  setDT(arms_x3)
  arms3_select <- arms_x3[obs.times %in% times_for_tables]
  abs_rates <- dcast(arms3_select, obs.times ~ arm, value.var="n1_cum_per")
  abs_rates[,pl:= round(100*pl,1)][, tx:= round(100*tx,1)]
  colnames(abs_rates) <- c("Time (days)", "Events in untreated (%)", "Events in treated (%)")
  
  arr_and_individ <- merge(abs_rates, arr_cis_select, by="Time (days)")
  ft<- flextable(arr_and_individ)
  ft<- width(ft,width=dim_pretty(ft)$widths)
  
  cat(paste0('##### ', strata_y[i], '\n\n'))
  
  flextable_to_rmd(ft)
  
  cat("\n\n")
}
```
##### Overall

```{r arr in one table, results='asis'}

#calculate ARR for whole cohort
#matrix of risk ratios at t1, one column per stratum, nrep rows for each day. Need to do this once for each day.
arr_per_strata <- unlist(lapply(arr_2, function(x) x[,V3, drop=FALSE]))

arr_flat<-matrix(arr_per_strata, ncol=length(strata_y))

#matrix of ns in latch, one column per stratum
strata_n_df<- as.data.frame(t(strata_y_vars$strata_y_n))
ns <- as.matrix(strata_n_df %>% slice(rep(1, each = nrep*length(all_times))))
#create a vector of totals
n_tot<- rowSums(ns)

arr_n <- rowSums(arr_flat*ns)
arr_n_tot <- as.data.frame(arr_n/n_tot)

setDT(arr_n_tot)
arr_n_tot[, index:=rep(all_times, each=nrep)]

#Group by index then work out quantiles. Use data.table
arr_n_tot[,mean_both:= (mean(`arr_n/n_tot`,na.rm=T)*100), by=index][,l_ci_both:=(quantile(`arr_n/n_tot`, probs=0.025, na.rm=T)*100),by=index][, u_ci_both:=(quantile(`arr_n/n_tot`, probs=0.975,na.rm=T)*100), by=index]

#select every nrep row
arr_n_tot_new <- arr_n_tot[seq(1, nrow(arr_n_tot),nrep), ]

#a note on the formatting of the risks outputs. For ease of calculating the confidence intervals (to keep compatible with the function get_CIs), the data frames include unnecessary columns. We select the relevant ones now.

arr_rel_cols<- list()
for(i in 1: length(strata_y)){
setDT(arr_cis[[i]])
arr_rel_cols[[i]]<-  arr_cis[[i]][,c("time", "mean_both", "l_ci_both", "u_ci_both")][, strata:=strata_y[i]][, mean_both:= 100*mean_both][, l_ci_both:= 100*l_ci_both] [,u_ci_both:= 100*u_ci_both]
}

arr_all <- do.call(rbind.data.frame, arr_rel_cols)
#we want to add in rows for the combined weighted total. First harmonise column names

arr_comb_1 <- arr_n_tot_new[, c("index","mean_both","l_ci_both","u_ci_both")][,strata:="All"]
setnames(arr_comb_1, "index", "time")

arr_all_comb<- rbind(arr_all, arr_comb_1)

### table at various time points
arr_select <- arr_n_tot_new[index %in% times_for_tables][, `arr_n/n_tot`:=NULL]
colnames(arr_select)<- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)")

ft_comb<- flextable(round(arr_select,1))
ft_comb<- width(ft_comb,width=dim_pretty(ft_comb)$widths)

ft_comb
```

#### Plots {.tabset}


##### Plot of ARR for each strata

```{r plot overall, results='asis', fig.height=4, fig.width=9}

#all on one plot
p_real <- ggplot(data=arr_all)+
  geom_smooth( aes(x =time, y = mean_both, colour=strata), linewidth=0.5, linetype="dashed")+
  geom_smooth( data=arr_comb_1,aes(x =time, y = mean_both, colour=strata), linewidth=1)+
  geom_ribbon(data=arr_comb_1, aes(ymin = l_ci_both, ymax = u_ci_both, x=time),fill= "#377EB8",alpha=0.1)+ 
  scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
               labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
  scale_y_continuous ("Absolute risk reduction (%)") +
  scale_color_discrete(palette=c("#377EB8","#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+
  theme_bw()+
  ggtitle(paste("Absolute risk reduction of recurrent stroke, accounting for competing risk of death"))+
  labs(color="Stratum")
p_real
ggsave("ARR plots with varying competing risks/Real competing risk ratios, combined.png")

```

##### Plot of ARR for each stratum separately

```{r plot sep, fig.height=4, fig.width=12}
#strata separately
ggplot(data=arr_all)+
  geom_smooth( aes(x =time, y = mean_both,color=strata), linewidth=0.5, linetype="dashed")+
  geom_ribbon( aes(ymin = l_ci_both, ymax = u_ci_both, x=time,fill=strata),alpha=0.1)+ 
  facet_wrap(facets="strata", ncol=4, nrow=2)+
  scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6),
               labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) +
  scale_y_continuous ("Absolute risk reduction (%)") +
  scale_color_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+
  scale_fill_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+
  theme_bw()+
  ggtitle(paste("Absolute risk reduction of recurrent stroke, accounting for competing risk of death"))+
  labs(color="Stratum")

ggsave("ARR plots with varying competing risks/Real competing risk ratios, separate.png")
```


## Appendix
  
  We pool the risk ratio at $t_1$, to find the calibrated relative risk after 18 months.

```{r t1 pooled results printout, comment=NA}
cat(paste("The estimated calibrated risk ratio is ", est_t1, " (95% confidence interval ", cis_t1[1], "-", cis_t1[2], ")", sep=""))
```

\n
\n

<!-- ```{r t1 pooled results,results="asis",fig.width=3,fig.height=2} -->
<!-- ggplot(data=as.data.frame(rr_tot_t1))+ -->
<!--   geom_histogram(aes(x=rr_tot_t1), fill="#FCCDE5")+ -->
<!--   theme_bw()+ -->
<!--   labs(title="Simulated risk ratio, T1")+ -->
<!--   xlab("Risk ratio") -->

<!-- ``` -->


<!-- ## Sensitivity analysis -->

<!-- ## Sensitivity analyses -->


<!-- The confidence intervals were wide for the competing risk, so may not be valid. We therefore rerun with three different competing risk hazard ratios: 1; 0.9 (95%CI 0.7-1.1); and 0.8 (0.6-1.8). -->


<!-- ### Small difference in competing risk (HR for competing risk = 0.9 (0.7-1.1)) -->

<!-- The confidence intervals were wide for the competing risk, so may not be valid. We therefore rerun with no difference in competing risk between each arm.  -->


<!-- ```{r small comp risk difference} -->

<!-- # calculate the log hr, and log se from normal distribution formulae -->

<!-- HR_comp_t1<- data.frame(HR=0.9, l_ci=0.7, u_ci=1.1) -->
<!-- HR_comp_t1<- HR_comp_t1[rep(seq_len(nrow(HR_comp_t1)), each=8),] -->
<!-- HR_comp_t2<- copy(HR_comp_t1) -->
<!-- setDT(HR_comp_t1) -->
<!-- setDT(HR_comp_t2) -->

<!-- HR_comp_t1[,log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)] -->
<!-- HR_comp_t2[,log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)] -->

<!-- # simulate nrep samples. .I means, do each row nrep times. The output is nrep rows for each stratum, so rows 1:nrep are stratum 1, (nrep+1:2*nrep) are stratum 2, etc. This can be confirmed by checking the mean of each chunk of rows.  -->
<!-- log_hr_comp_t1<- HR_comp_t1[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I] -->
<!-- log_hr_comp_t2<- HR_comp_t2[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I] -->

<!-- #for each time period, generate a rate of competing risk in the treated. -->
<!-- log_lambda_t_comp_t1 <- copy(log_lambda_u_comp) -->
<!-- log_lambda_t_comp_t2 <- copy(log_lambda_u_comp) -->
<!-- log_lambda_t_comp_t1[, log_rate_comp_t_est_t1:=V1+log_hr_comp_t1$V1 ][,V1:=NULL] -->
<!-- log_lambda_t_comp_t2[, log_rate_comp_t_est_t2:=V1+log_hr_comp_t2$V1 ][,V1:=NULL] -->
<!-- #harmonise names with untreated -->
<!-- setnames(log_lambda_t_comp_t1, "log_rate_comp_t_est_t1", "V1") -->
<!-- setnames(log_lambda_t_comp_t2, "log_rate_comp_t_est_t2", "V1") -->

<!-- #calculate confidence intervals for visualisation -->
<!-- conf_int_treated_t1 <- log_lambda_t_comp_t1[,quantile(V1, c(0.025,0.5,0.975)), by=I] -->
<!-- conf_int_treated_t1[, cat:= rep(c("Lower CI", "Mean", "Upper CI"), length(strata_y))] -->

<!-- conf_int_treated_t2 <- log_lambda_t_comp_t2[,quantile(V1, c(0.025,0.5,0.975)), by=I] -->
<!-- conf_int_treated_t2[, cat:= rep(c("Lower CI", "Mean", "Upper CI"), length(strata_y))] -->

<!-- #widen table -->
<!-- treated_comp_t1 <-dcast(conf_int_treated_t1, I~cat, value.var = "V1") -->
<!-- treated_comp_t1_x<- treated_comp_t1[, "Rate per 100 person/years":=paste(round(exp(Mean)*100*365.25,2), " (",round(exp(`Lower CI`)*100*365.25,2), " - ",round(exp(`Upper CI`)*100*365.25,2),")", sep="") ][, "Lower CI":=NULL][, "Upper CI":=NULL] -->
<!-- colnames(treated_comp_t1_x) <- c( "Stratum",  "log(rate/day)","Events per 100 person years (95% CI)") -->

<!-- treated_comp_t2 <-dcast(conf_int_treated_t2, I~cat, value.var = "V1") -->
<!-- treated_comp_t2_x<- treated_comp_t2[, "Rate per 100 person/years":=paste(round(exp(Mean)*100*365.25,2), " (",round(exp(`Lower CI`)*100*365.25,2), " - ",round(exp(`Upper CI`)*100*365.25,2),")", sep="") ][, "Lower CI":=NULL][, "Upper CI":=NULL] -->
<!-- colnames(treated_comp_t2_x) <- c( "Stratum",  "log(rate/day)","Events per 100 person years (95% CI)") -->

<!-- # table of rates combined -->
<!-- setDT(untreated_rates_comp_round) -->
<!-- unt_for_merge <- untreated_rates_comp_round[, Stratum:=strata_y][, c("Stratum", "rate_cis")] -->
<!-- tr_comp_t1_for_m<- treated_comp_t1_x[, c("Stratum", "Events per 100 person years (95% CI)")] -->
<!-- tr_comp_t2_for_m<- treated_comp_t2_x[, c("Stratum", "Events per 100 person years (95% CI)")] -->
<!-- setnames(unt_for_merge, "rate_cis", "Events per 100py, untreated") -->
<!-- setnames(tr_comp_t1_for_m, "Events per 100 person years (95% CI)", "Events per 100py, treated t1") -->
<!-- setnames(tr_comp_t2_for_m, "Events per 100 person years (95% CI)", "Events per 100py, treated t2") -->
<!-- m_1<- merge(unt_for_merge, tr_comp_t1_for_m, by="Stratum") -->
<!-- rates_comp_comb<- merge(m_1, tr_comp_t2_for_m, by="Stratum") -->
<!-- rates_comp_comb$Stratum<- factor(rates_comp_comb$Stratum, levels=strata_y, ordered=T) -->
<!-- rates_comp_comb<- rates_comp_comb%>% arrange(Stratum) -->
<!-- rm(m_1, unt_for_merge, tr_comp_t1_for_m, tr_comp_t2_for_m) -->



<!-- ``` -->

<!-- ```{r loop through strata for arr comp small, results='asis'} -->

<!-- #We want arr for each stratum. The lambdas are in a single vector, length of the number of strata time the number of reps. -->
<!-- pl_2 <- list() -->
<!-- pl_2_cis <- list() -->
<!-- tx_2 <- list() -->
<!-- tx_2_cis <- list() -->

<!-- arr_2 <- list() -->
<!-- rr_2 <- list() -->
<!-- or_2<- list() -->

<!-- for(i in 1:length(strata_y)){ -->
<!--   # Placebo arm -->
<!--   pl_2[[i]] <- MakeCumulative2(arm="u", -->
<!--                          rate.main.t1 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                          rate.main.t2 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                          rate.compete.t1 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                          rate.compete.t2 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                          t1_times, -->
<!--                          t2_times) -->
<!--   pl_2_cis[[i]]<- get_CIs(pl_2[[i]], c(t1_times, t2_times)) -->

<!--   # Treatment arm -->
<!--   tx_2[[i]] <- MakeCumulative2 (arm = "rx", -->
<!--                           rate.main.t1 = lambda_t1_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                           rate.main.t2 = lambda_t2_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                           rate.compete.t1 = exp(log_lambda_t_comp_t1[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                           rate.compete.t2 = exp(log_lambda_t_comp_t2[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                           t1_times, -->
<!--                           t2_times) -->
<!--   tx_2_cis[[i]]<- get_CIs(tx_2[[i]],c(t1_times, t2_times)) -->
<!--   ## Calculate comparisons -->
<!--   arr_2[[i]] = pl_2[[i]] - tx_2[[i]] -->
<!--   rr_2[[i]] = tx_2[[i]]/pl_2[[i]] -->
<!--   or_2[[i]] = (tx_2[[i]]/(1-tx_2[[i]])) / (pl_2[[i]]/(1-pl_2[[i]])) -->

<!-- } -->


<!-- ``` -->


<!-- ```{r plots of outcomes comp small, fig.width=12, fig.height=6, results='hide'} -->
<!-- plots_rec_str<- list() -->
<!-- arms_x1<- list() -->
<!-- for(i in 1:length(strata_y)){ -->

<!-- arms_x1[[i]] <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_main, l_ci = pl_2_cis[[i]]$l_ci_main, u_ci=pl_2_cis[[i]]$u_ci_main), -->
<!--                     tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_main, l_ci = tx_2_cis[[i]]$l_ci_main, u_ci=tx_2_cis[[i]]$u_ci_main)) -->

<!-- ## Create plots -->
<!-- setDT(arms_x1[[i]]) -->
<!-- plots_rec_str[[i]] <- ggplot(arms_x1[[i]], aes(x = obs.times, y = n1_cum_per, colour = arm)) + -->
<!-- geom_ribbon(aes(ymin = l_ci, ymax = u_ci, colour=NULL, fill=arm), alpha=0.2) + -->
<!-- geom_line()+ -->
<!-- #geom_point()+ -->
<!-- scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!-- scale_y_continuous ("Cumulative incidence", limits=c(0,0.5)) + -->
<!-- theme_bw()+ -->
<!-- theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+ -->
<!-- ggtitle(paste(strata_y[i]))+ -->
<!-- guides(color="none")+ -->
<!-- labs(fill="Arm") -->
<!-- } -->

<!-- #(plots_rec_str[[1]]|plots_rec_str[[2]]|plots_rec_str[[3]]|plots_rec_str[[4]])/(plots_rec_str[[5]]|plots_rec_str[[6]]|plots_rec_str[[7]]|plots_rec_str[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke for each stratum") -->

<!-- ``` -->


<!-- ```{r competing outcome plots comp small,fig.width=12, fig.height=6, results='hide'} -->
<!-- plots_comp = list() -->
<!-- for(i in 1:length(strata_y)){ -->

<!--     arms_x2 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_comp, l_ci = pl_2_cis[[i]]$l_ci_comp, u_ci=pl_2_cis[[i]]$u_ci_comp), -->
<!--                    tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_comp, l_ci = tx_2_cis[[i]]$l_ci_comp, u_ci=tx_2_cis[[i]]$u_ci_comp)) -->
<!--     ## set upper limit on upper ci of 1 so it plots properly -->

<!--     arms_x2<- arms_x2%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci)) -->
<!--     ## Create plots -->
<!--     plots_comp[[i]] <- ggplot(arms_x2, aes(x = obs.times, y = n1_cum_per, colour = arm)) + -->
<!--     geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) + -->
<!--     geom_line()+ -->
<!--     scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                    labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!--     scale_y_continuous ("Cumulative incidence", limits=c(0,1)) + -->
<!--     theme_bw()+ -->
<!--     theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+ -->
<!--     ggtitle(paste(strata_y[i]))+ -->
<!--     guides(color="none")+ -->
<!--     labs(fill="Arm") -->
<!-- } -->

<!-- #(plots_comp[[1]]|plots_comp[[2]]|plots_comp[[3]]|plots_comp[[4]])/(plots_comp[[5]]|plots_comp[[6]]|plots_comp[[7]]|plots_comp[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of competing risk for each stratum") -->

<!-- ``` -->


<!-- ```{r both outcome plots comp small,fig.width=12, fig.height=6,  results='hide'} -->

<!-- plot_both= list() -->
<!-- for(i in 1:length(strata_y)){ -->

<!--     arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both), -->
<!--                    tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both)) -->


<!--     ## set upper limit on upper ci of 1 so it plots properly -->

<!--     arms_x3<- arms_x3%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci)) -->
<!--     ## Create plots -->
<!--     plot_both[[i]] <- ggplot(arms_x3, aes(x = obs.times, y = n1_cum_per, colour = arm)) + -->
<!--     geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) + -->
<!--     geom_line()+ -->
<!--     scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                    labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!--     scale_y_continuous ("Cumulative incidence", limits=c(0,1)) + -->
<!--     theme_bw()+ -->
<!--     theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+ -->
<!--     ggtitle(paste(strata_y[i]))+ -->
<!--     guides(color="none")+ -->
<!--     labs(fill="Arm") -->

<!-- } -->

<!-- #(plot_both[[1]]|plot_both[[2]]|plot_both[[3]]|plot_both[[4]])/(plot_both[[5]]|plot_both[[6]]|plot_both[[7]]|plot_both[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke, accounting for competing risk of death") -->

<!-- ``` -->

<!-- #### Absolute risk reduction {.tabset} -->

<!-- These plots and tables show the absolute risk reduction for each stratum and overall, accounting for competing risk of death. -->

<!-- ```{r format ARR comp small, results='asis'} -->
<!-- #create dataframes and tables of arr and event rates in each arm, for each stratum -->
<!-- arr_cis<- list()  -->
<!-- for(i in 1:length(strata_y)){ -->

<!-- arr_cis[[i]] <- as.data.frame(get_CIs(arr_2[[i]], all_times)) -->
<!-- setDT(arr_cis[[i]]) -->
<!-- arr_cis[[i]][, time:= all_times] -->
<!-- arr_cis_select<- arr_cis[[i]][time %in% times_for_tables][,c("time","mean_both", "l_ci_both", "u_ci_both") ] -->
<!-- arr_cis_select[, mean_both:= round(mean_both*100,1)][, u_ci_both:=round(u_ci_both*100,1)][, l_ci_both:=round(l_ci_both*100,1)] -->

<!-- colnames(arr_cis_select) <- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)") -->

<!-- arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "pl", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both), -->
<!--                tibble(obs.times = c(t1_times, t2_times), arm = "tx", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both)) -->

<!-- setDT(arms_x3) -->
<!-- arms3_select <- arms_x3[obs.times %in% times_for_tables] -->
<!-- abs_rates <- dcast(arms3_select, obs.times ~ arm, value.var="n1_cum_per") -->
<!-- abs_rates[,pl:= round(100*pl,1)][, tx:= round(100*tx,1)] -->
<!-- colnames(abs_rates) <- c("Time (days)", "Events in untreated (%)", "Events in treated (%)") -->

<!-- arr_and_individ <- merge(abs_rates, arr_cis_select, by="Time (days)") -->
<!-- ft<- flextable(arr_and_individ) -->
<!-- ft<- width(ft,width=dim_pretty(ft)$widths) -->

<!-- cat(paste0('##### ', strata_y[i], '\n\n')) -->

<!-- flextable_to_rmd(ft) -->

<!-- cat("\n\n") -->
<!-- } -->
<!-- ``` -->

<!-- ##### Overall -->

<!-- ```{r arr in one table comp small, results='asis'} -->

<!-- #calculate ARR for whole cohort -->
<!-- #matrix of risk ratios at t1, one column per stratum, nrep rows for each day. Need to do this once for each day. -->
<!-- arr_per_strata <- unlist(lapply(arr_2, function(x) x[,V3, drop=FALSE])) -->

<!-- arr_flat<-matrix(arr_per_strata, ncol=length(strata_y)) -->

<!-- #matrix of ns in latch, one column per stratum -->
<!-- strata_n_df<- as.data.frame(t(strata_y_vars$strata_y_n)) -->
<!-- ns <- as.matrix(strata_n_df %>% slice(rep(1, each = nrep*length(all_times)))) -->
<!-- #create a vector of totals -->
<!-- n_tot<- rowSums(ns) -->

<!-- arr_n <- rowSums(arr_flat*ns) -->
<!-- arr_n_tot <- as.data.frame(arr_n/n_tot) -->

<!-- setDT(arr_n_tot) -->
<!-- arr_n_tot[, index:=rep(all_times, each=nrep)] -->

<!-- #Group by index then work out quantiles. Use data.table -->
<!-- arr_n_tot[,mean_both:= (mean(`arr_n/n_tot`,na.rm=T)*100), by=index][,l_ci_both:=(quantile(`arr_n/n_tot`, probs=0.025, na.rm=T)*100),by=index][, u_ci_both:=(quantile(`arr_n/n_tot`, probs=0.975,na.rm=T)*100), by=index] -->

<!-- #select every nrep row -->
<!-- arr_n_tot_new <- arr_n_tot[seq(1, nrow(arr_n_tot),nrep), ] -->

<!-- #a note on the formatting of the risks outputs. For ease of calculating the confidence intervals (to keep compatible with the function get_CIs), the data frames include unnecessary columns. We select the relevant ones now. -->

<!-- arr_rel_cols<- list() -->
<!-- for(i in 1: length(strata_y)){ -->
<!-- setDT(arr_cis[[i]]) -->
<!-- arr_rel_cols[[i]]<-  arr_cis[[i]][,c("time", "mean_both", "l_ci_both", "u_ci_both")][, strata:=strata_y[i]][, mean_both:= 100*mean_both][, l_ci_both:= 100*l_ci_both] [,u_ci_both:= 100*u_ci_both] -->
<!-- } -->

<!-- arr_all <- do.call(rbind.data.frame, arr_rel_cols) -->
<!-- #we want to add in rows for the combined weighted total. First harmonise column names -->

<!-- arr_comb_1 <- arr_n_tot_new[, c("index","mean_both","l_ci_both","u_ci_both")][,strata:="All"] -->
<!-- setnames(arr_comb_1, "index", "time") -->

<!-- arr_all_comb<- rbind(arr_all, arr_comb_1) -->

<!-- ### table at various time points -->
<!-- arr_select <- arr_n_tot_new[index %in% times_for_tables][, `arr_n/n_tot`:=NULL] -->
<!-- colnames(arr_select)<- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)") -->

<!-- ft_comb<- flextable(round(arr_select,1)) -->
<!-- ft_comb<- width(ft_comb,width=dim_pretty(ft_comb)$widths) -->

<!-- ft_comb -->
<!-- ``` -->

<!-- #### Plots {.tabset} -->

<!-- ##### Plot of ARR for each strata -->

<!-- Absolute risk reduction for the whole LATCH population and each age/sex stratum, taking into account competing risk of death. -->
<!-- ```{r plot sep comp small, fig.height=4, fig.width=12} -->
<!-- #strata separately -->
<!--     ggplot(data=arr_all)+ -->
<!--         geom_smooth( aes(x =time, y = mean_both,color=strata), linewidth=0.5, linetype="dashed")+ -->
<!--         geom_ribbon( aes(ymin = l_ci_both, ymax = u_ci_both, x=time,fill=strata),alpha=0.1)+  -->
<!--         facet_wrap(facets="strata", ncol=4, nrow=2)+ -->
<!--         scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                      labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!--         scale_y_continuous ("Absolute risk reduction (%)") + -->
<!--         scale_color_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+ -->
<!--         scale_fill_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+ -->
<!--         theme_bw()+ -->
<!--         ggtitle(paste("Absolute risk reduction of recurrent stroke, HR for competing risk of death = 0.9"))+ -->
<!--         labs(color="Stratum") -->
<!--     ggsave("ARR plots with varying competing risks/Competing risk ratio 0.9, separate.png") -->
<!-- ``` -->

<!-- ##### Plot of ARR for each strata -->

<!-- ```{r plot overall comp small, results='asis', fig.height=4, fig.width=9} -->
<!-- #all on one plot -->
<!-- p_9<- ggplot(data=arr_all)+ -->
<!--     geom_smooth( aes(x =time, y = mean_both, colour=strata), linewidth=0.5, linetype="dashed")+ -->
<!--     geom_smooth( data=arr_comb_1,aes(x =time, y = mean_both, colour=strata), linewidth=1)+ -->
<!--     geom_ribbon(data=arr_comb_1, aes(ymin = l_ci_both, ymax = u_ci_both, x=time),fill= "#377EB8",alpha=0.1)+  -->
<!--     scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                  labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!--     scale_y_continuous ("Absolute risk reduction (%)") + -->
<!--     scale_color_discrete(palette=c("#377EB8","#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+ -->
<!--     theme_bw()+ -->
<!--     ggtitle(paste("Absolute risk reduction of recurrent stroke,HR for competing risk of death = 0.9"))+ -->
<!--     labs(color="Stratum") -->
<!-- p_9 -->
<!-- ggsave("ARR plots with varying competing risks/Competing risk ratio 0.9, combined.png") -->
<!-- ``` -->


<!-- ### Small difference in competing risk (HR for competing risk = 0.8 (0.6-1.0)) -->

<!-- The confidence intervals were wide for the competing risk, so may not be valid. We therefore rerun with no difference in competing risk between each arm.  -->


<!-- ```{r small comp risk difference 8} -->

<!-- # calculate the log hr, and log se from normal distribution formulae -->

<!-- HR_comp_t1<- data.frame(HR=0.8, l_ci=0.6, u_ci=1.0) -->
<!-- HR_comp_t1<- HR_comp_t1[rep(seq_len(nrow(HR_comp_t1)), each=8),] -->
<!-- HR_comp_t2<- copy(HR_comp_t1) -->
<!-- setDT(HR_comp_t1) -->
<!-- setDT(HR_comp_t2) -->

<!-- HR_comp_t1[,log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)] -->
<!-- HR_comp_t2[,log_hr:= log(HR)][, log_se:= (log(u_ci)-log(l_ci))/3.92][, se:= exp(log_se)] -->

<!-- # simulate nrep samples. .I means, do each row nrep times. The output is nrep rows for each stratum, so rows 1:nrep are stratum 1, (nrep+1:2*nrep) are stratum 2, etc. This can be confirmed by checking the mean of each chunk of rows.  -->
<!-- log_hr_comp_t1<- HR_comp_t1[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I] -->
<!-- log_hr_comp_t2<- HR_comp_t2[,rnorm(nrep, mean=log_hr, sd=log_se), by=.I] -->

<!-- #for each time period, generate a rate of competing risk in the treated. -->
<!-- log_lambda_t_comp_t1 <- copy(log_lambda_u_comp) -->
<!-- log_lambda_t_comp_t2 <- copy(log_lambda_u_comp) -->
<!-- log_lambda_t_comp_t1[, log_rate_comp_t_est_t1:=V1+log_hr_comp_t1$V1 ][,V1:=NULL] -->
<!-- log_lambda_t_comp_t2[, log_rate_comp_t_est_t2:=V1+log_hr_comp_t2$V1 ][,V1:=NULL] -->
<!-- #harmonise names with untreated -->
<!-- setnames(log_lambda_t_comp_t1, "log_rate_comp_t_est_t1", "V1") -->
<!-- setnames(log_lambda_t_comp_t2, "log_rate_comp_t_est_t2", "V1") -->

<!-- #calculate confidence intervals for visualisation -->
<!-- conf_int_treated_t1 <- log_lambda_t_comp_t1[,quantile(V1, c(0.025,0.5,0.975)), by=I] -->
<!-- conf_int_treated_t1[, cat:= rep(c("Lower CI", "Mean", "Upper CI"), length(strata_y))] -->

<!-- conf_int_treated_t2 <- log_lambda_t_comp_t2[,quantile(V1, c(0.025,0.5,0.975)), by=I] -->
<!-- conf_int_treated_t2[, cat:= rep(c("Lower CI", "Mean", "Upper CI"), length(strata_y))] -->

<!-- #widen table -->
<!-- treated_comp_t1 <-dcast(conf_int_treated_t1, I~cat, value.var = "V1") -->
<!-- treated_comp_t1_x<- treated_comp_t1[, "Rate per 100 person/years":=paste(round(exp(Mean)*100*365.25,2), " (",round(exp(`Lower CI`)*100*365.25,2), " - ",round(exp(`Upper CI`)*100*365.25,2),")", sep="") ][, "Lower CI":=NULL][, "Upper CI":=NULL] -->
<!-- colnames(treated_comp_t1_x) <- c( "Stratum",  "log(rate/day)","Events per 100 person years (95% CI)") -->

<!-- treated_comp_t2 <-dcast(conf_int_treated_t2, I~cat, value.var = "V1") -->
<!-- treated_comp_t2_x<- treated_comp_t2[, "Rate per 100 person/years":=paste(round(exp(Mean)*100*365.25,2), " (",round(exp(`Lower CI`)*100*365.25,2), " - ",round(exp(`Upper CI`)*100*365.25,2),")", sep="") ][, "Lower CI":=NULL][, "Upper CI":=NULL] -->
<!-- colnames(treated_comp_t2_x) <- c( "Stratum",  "log(rate/day)","Events per 100 person years (95% CI)") -->

<!-- # table of rates combined -->
<!-- setDT(untreated_rates_comp_round) -->
<!-- unt_for_merge <- untreated_rates_comp_round[, Stratum:=strata_y][, c("Stratum", "rate_cis")] -->
<!-- tr_comp_t1_for_m<- treated_comp_t1_x[, c("Stratum", "Events per 100 person years (95% CI)")] -->
<!-- tr_comp_t2_for_m<- treated_comp_t2_x[, c("Stratum", "Events per 100 person years (95% CI)")] -->
<!-- setnames(unt_for_merge, "rate_cis", "Events per 100py, untreated") -->
<!-- setnames(tr_comp_t1_for_m, "Events per 100 person years (95% CI)", "Events per 100py, treated t1") -->
<!-- setnames(tr_comp_t2_for_m, "Events per 100 person years (95% CI)", "Events per 100py, treated t2") -->
<!-- m_1<- merge(unt_for_merge, tr_comp_t1_for_m, by="Stratum") -->
<!-- rates_comp_comb<- merge(m_1, tr_comp_t2_for_m, by="Stratum") -->
<!-- rates_comp_comb$Stratum<- factor(rates_comp_comb$Stratum, levels=strata_y, ordered=T) -->
<!-- rates_comp_comb<- rates_comp_comb%>% arrange(Stratum) -->
<!-- rm(m_1, unt_for_merge, tr_comp_t1_for_m, tr_comp_t2_for_m) -->



<!-- ``` -->

<!-- ```{r loop through strata for arr comp 8, results='asis'} -->

<!-- #We want arr for each stratum. The lambdas are in a single vector, length of the number of strata time the number of reps. -->
<!-- pl_2 <- list() -->
<!-- pl_2_cis <- list() -->
<!-- tx_2 <- list() -->
<!-- tx_2_cis <- list() -->

<!-- arr_2 <- list() -->
<!-- rr_2 <- list() -->
<!-- or_2<- list() -->

<!-- for(i in 1:length(strata_y)){ -->
<!--     # Placebo arm -->
<!--     pl_2[[i]] <- MakeCumulative2(arm="u", -->
<!--                            rate.main.t1 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                            rate.main.t2 = lambda_u_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                            rate.compete.t1 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                            rate.compete.t2 = exp(log_lambda_u_comp[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                            t1_times, -->
<!--                            t2_times) -->
<!--     pl_2_cis[[i]]<- get_CIs(pl_2[[i]], c(t1_times, t2_times)) -->

<!--     # Treatment arm -->
<!--     tx_2[[i]] <- MakeCumulative2 (arm = "rx", -->
<!--                             rate.main.t1 = lambda_t1_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                             rate.main.t2 = lambda_t2_sim[(nrep*(i-1)+1):(i*nrep)], -->
<!--                             rate.compete.t1 = exp(log_lambda_t_comp_t1[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                             rate.compete.t2 = exp(log_lambda_t_comp_t2[(nrep*(i-1)+1):(i*nrep),2]), -->
<!--                             t1_times, -->
<!--                             t2_times) -->
<!--     tx_2_cis[[i]]<- get_CIs(tx_2[[i]],c(t1_times, t2_times)) -->
<!--     ## Calculate comparisons -->
<!--     arr_2[[i]] = pl_2[[i]] - tx_2[[i]] -->
<!--     rr_2[[i]] = tx_2[[i]]/pl_2[[i]] -->
<!--     or_2[[i]] = (tx_2[[i]]/(1-tx_2[[i]])) / (pl_2[[i]]/(1-pl_2[[i]])) -->

<!-- } -->


<!-- ``` -->


<!-- ```{r plots of outcomes comp 8, fig.width=12, fig.height=6, results='hide'} -->
<!-- plots_rec_str<- list() -->
<!-- arms_x1<- list() -->
<!-- for(i in 1:length(strata_y)){ -->

<!-- arms_x1[[i]] <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_main, l_ci = pl_2_cis[[i]]$l_ci_main, u_ci=pl_2_cis[[i]]$u_ci_main), -->
<!--                     tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_main, l_ci = tx_2_cis[[i]]$l_ci_main, u_ci=tx_2_cis[[i]]$u_ci_main)) -->

<!-- ## Create plots -->
<!-- setDT(arms_x1[[i]]) -->
<!-- plots_rec_str[[i]] <- ggplot(arms_x1[[i]], aes(x = obs.times, y = n1_cum_per, colour = arm)) + -->
<!-- geom_ribbon(aes(ymin = l_ci, ymax = u_ci, colour=NULL, fill=arm), alpha=0.2) + -->
<!-- geom_line()+ -->
<!-- #geom_point()+ -->
<!-- scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!-- scale_y_continuous ("Cumulative incidence", limits=c(0,0.5)) + -->
<!-- theme_bw()+ -->
<!-- theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+ -->
<!-- ggtitle(paste(strata_y[i]))+ -->
<!-- guides(color="none")+ -->
<!-- labs(fill="Arm") -->
<!-- } -->

<!-- #(plots_rec_str[[1]]|plots_rec_str[[2]]|plots_rec_str[[3]]|plots_rec_str[[4]])/(plots_rec_str[[5]]|plots_rec_str[[6]]|plots_rec_str[[7]]|plots_rec_str[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke for each stratum") -->

<!-- ``` -->


<!-- ```{r competing outcome plots comp 8,fig.width=12, fig.height=6, results='hide'} -->
<!-- plots_comp = list() -->
<!-- for(i in 1:length(strata_y)){ -->

<!--     arms_x2 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_comp, l_ci = pl_2_cis[[i]]$l_ci_comp, u_ci=pl_2_cis[[i]]$u_ci_comp), -->
<!--                    tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_comp, l_ci = tx_2_cis[[i]]$l_ci_comp, u_ci=tx_2_cis[[i]]$u_ci_comp)) -->
<!--     ## set upper limit on upper ci of 1 so it plots properly -->

<!--     arms_x2<- arms_x2%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci)) -->
<!--     ## Create plots -->
<!--     plots_comp[[i]] <- ggplot(arms_x2, aes(x = obs.times, y = n1_cum_per, colour = arm)) + -->
<!--     geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) + -->
<!--     geom_line()+ -->
<!--     scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                    labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!--     scale_y_continuous ("Cumulative incidence", limits=c(0,1)) + -->
<!--     theme_bw()+ -->
<!--     theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+ -->
<!--     ggtitle(paste(strata_y[i]))+ -->
<!--     guides(color="none")+ -->
<!--     labs(fill="Arm") -->
<!-- } -->

<!-- #(plots_comp[[1]]|plots_comp[[2]]|plots_comp[[3]]|plots_comp[[4]])/(plots_comp[[5]]|plots_comp[[6]]|plots_comp[[7]]|plots_comp[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of competing risk for each stratum") -->

<!-- ``` -->


<!-- ```{r both outcome plots comp 8,fig.width=12, fig.height=6,  results='hide'} -->

<!-- plot_both= list() -->
<!-- for(i in 1:length(strata_y)){ -->

<!-- arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "Placebo", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both), -->
<!--                tibble(obs.times = c(t1_times, t2_times), arm = "Treatment", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both)) -->


<!-- ## set upper limit on upper ci of 1 so it plots properly -->

<!-- arms_x3<- arms_x3%>% mutate(u_ci_mod = ifelse(u_ci>1,1,u_ci)) -->
<!-- ## Create plots -->
<!-- plot_both[[i]] <- ggplot(arms_x3, aes(x = obs.times, y = n1_cum_per, colour = arm)) + -->
<!-- geom_ribbon(aes(ymin = l_ci, ymax = u_ci_mod, colour=NULL, fill=arm), alpha=0.2) + -->
<!-- geom_line()+ -->
<!-- scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!-- scale_y_continuous ("Cumulative incidence", limits=c(0,1)) + -->
<!-- theme_bw()+ -->
<!-- theme(plot.title=element_text(size=8), axis.title = element_text(size=8), axis.text=element_text(size=8), legend.text = element_text(size=8), legend.title=element_text(size=8))+ -->
<!-- ggtitle(paste(strata_y[i]))+ -->
<!-- guides(color="none")+ -->
<!-- labs(fill="Arm") -->

<!-- } -->

<!-- #(plot_both[[1]]|plot_both[[2]]|plot_both[[3]]|plot_both[[4]])/(plot_both[[5]]|plot_both[[6]]|plot_both[[7]]|plot_both[[8]])+plot_layout(guides="collect", axis_titles = "collect")+plot_annotation(title= "Cumulative incidence of recurrent stroke, accounting for competing risk of death") -->

<!-- ``` -->

<!-- #### Absolute risk reduction {.tabset} -->

<!-- These plots and tables show the absolute risk reduction for each stratum and overall, accounting for competing risk of death. -->

<!-- ```{r format ARR comp 8, results='asis'} -->
<!-- #create dataframes and tables of arr and event rates in each arm, for each stratum -->
<!-- arr_cis<- list()  -->
<!-- for(i in 1:length(strata_y)){ -->

<!-- arr_cis[[i]] <- as.data.frame(get_CIs(arr_2[[i]], all_times)) -->
<!-- setDT(arr_cis[[i]]) -->
<!-- arr_cis[[i]][, time:= all_times] -->
<!-- arr_cis_select<- arr_cis[[i]][time %in% times_for_tables][,c("time","mean_both", "l_ci_both", "u_ci_both") ] -->
<!-- arr_cis_select[, mean_both:= round(mean_both*100,1)][, u_ci_both:=round(u_ci_both*100,1)][, l_ci_both:=round(l_ci_both*100,1)] -->

<!-- colnames(arr_cis_select) <- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)") -->

<!-- arms_x3 <- bind_rows(tibble(obs.times = c(t1_times, t2_times), arm = "pl", n1_cum_per = pl_2_cis[[i]]$mean_both, l_ci = pl_2_cis[[i]]$l_ci_both, u_ci=pl_2_cis[[i]]$u_ci_both), -->
<!--                tibble(obs.times = c(t1_times, t2_times), arm = "tx", n1_cum_per = tx_2_cis[[i]]$mean_both, l_ci = tx_2_cis[[i]]$l_ci_both, u_ci=tx_2_cis[[i]]$u_ci_both)) -->

<!-- setDT(arms_x3) -->
<!-- arms3_select <- arms_x3[obs.times %in% times_for_tables] -->
<!-- abs_rates <- dcast(arms3_select, obs.times ~ arm, value.var="n1_cum_per") -->
<!-- abs_rates[,pl:= round(100*pl,1)][, tx:= round(100*tx,1)] -->
<!-- colnames(abs_rates) <- c("Time (days)", "Events in untreated (%)", "Events in treated (%)") -->

<!-- arr_and_individ <- merge(abs_rates, arr_cis_select, by="Time (days)") -->
<!-- ft<- flextable(arr_and_individ) -->
<!-- ft<- width(ft,width=dim_pretty(ft)$widths) -->

<!-- cat(paste0('##### ', strata_y[i], '\n\n')) -->

<!-- flextable_to_rmd(ft) -->

<!-- cat("\n\n") -->
<!-- } -->
<!-- ``` -->

<!-- ##### Overall -->

<!-- ```{r arr in one table comp 8, results='asis'} -->

<!-- #calculate ARR for whole cohort -->
<!-- #matrix of risk ratios at t1, one column per stratum, nrep rows for each day. Need to do this once for each day. -->
<!-- arr_per_strata <- unlist(lapply(arr_2, function(x) x[,V3, drop=FALSE])) -->

<!-- arr_flat<-matrix(arr_per_strata, ncol=length(strata_y)) -->

<!-- #matrix of ns in latch, one column per stratum -->
<!-- strata_n_df<- as.data.frame(t(strata_y_vars$strata_y_n)) -->
<!-- ns <- as.matrix(strata_n_df %>% slice(rep(1, each = nrep*length(all_times)))) -->
<!-- #create a vector of totals -->
<!-- n_tot<- rowSums(ns) -->

<!-- arr_n <- rowSums(arr_flat*ns) -->
<!-- arr_n_tot <- as.data.frame(arr_n/n_tot) -->

<!-- setDT(arr_n_tot) -->
<!-- arr_n_tot[, index:=rep(all_times, each=nrep)] -->

<!-- #Group by index then work out quantiles. Use data.table -->
<!-- arr_n_tot[,mean_both:= (mean(`arr_n/n_tot`,na.rm=T)*100), by=index][,l_ci_both:=(quantile(`arr_n/n_tot`, probs=0.025, na.rm=T)*100),by=index][, u_ci_both:=(quantile(`arr_n/n_tot`, probs=0.975,na.rm=T)*100), by=index] -->

<!-- #select every nrep row -->
<!-- arr_n_tot_new <- arr_n_tot[seq(1, nrow(arr_n_tot),nrep), ] -->

<!-- #a note on the formatting of the risks outputs. For ease of calculating the confidence intervals (to keep compatible with the function get_CIs), the data frames include unnecessary columns. We select the relevant ones now. -->

<!-- arr_rel_cols<- list() -->
<!-- for(i in 1: length(strata_y)){ -->
<!-- setDT(arr_cis[[i]]) -->
<!-- arr_rel_cols[[i]]<-  arr_cis[[i]][,c("time", "mean_both", "l_ci_both", "u_ci_both")][, strata:=strata_y[i]][, mean_both:= 100*mean_both][, l_ci_both:= 100*l_ci_both] [,u_ci_both:= 100*u_ci_both] -->
<!-- } -->

<!-- arr_all <- do.call(rbind.data.frame, arr_rel_cols) -->
<!-- #we want to add in rows for the combined weighted total. First harmonise column names -->

<!-- arr_comb_1 <- arr_n_tot_new[, c("index","mean_both","l_ci_both","u_ci_both")][,strata:="All"] -->
<!-- setnames(arr_comb_1, "index", "time") -->

<!-- arr_all_comb<- rbind(arr_all, arr_comb_1) -->

<!-- ### table at various time points -->
<!-- arr_select <- arr_n_tot_new[index %in% times_for_tables][, `arr_n/n_tot`:=NULL] -->
<!-- colnames(arr_select)<- c("Time (days)", "Absolute difference (%)", "Lower CI (%)", "Upper CI (%)") -->

<!-- ft_comb<- flextable(round(arr_select,1)) -->
<!-- ft_comb<- width(ft_comb,width=dim_pretty(ft_comb)$widths) -->

<!-- ft_comb -->
<!-- ``` -->

<!-- #### Plots {.tabset} -->

<!-- ##### Plot of ARR for each strata -->

<!-- Absolute risk reduction for the whole LATCH population and each age/sex stratum, taking into account competing risk of death. -->
<!-- ```{r plot sep comp 8, fig.height=4, fig.width=12} -->
<!-- #strata separately -->
<!-- ggplot(data=arr_all)+ -->
<!-- geom_smooth( aes(x =time, y = mean_both,color=strata), linewidth=0.5, linetype="dashed")+ -->
<!-- geom_ribbon( aes(ymin = l_ci_both, ymax = u_ci_both, x=time,fill=strata),alpha=0.1)+  -->
<!-- facet_wrap(facets="strata", ncol=4, nrow=2)+ -->
<!-- scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--              labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!-- scale_y_continuous ("Absolute risk reduction (%)") + -->
<!-- scale_color_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+ -->
<!-- scale_fill_discrete(palette=c("#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+ -->
<!-- theme_bw()+ -->
<!-- ggtitle(paste("Absolute risk reduction of recurrent stroke, HR for competing risk of death = 0.8"))+ -->
<!-- labs(color="Stratum") -->
<!-- ggsave("ARR plots with varying competing risks/Competing risk ratio 0.8, separate.png") -->
<!-- ``` -->

<!-- ##### Plot of ARR for each strata -->

<!-- ```{r plot overall comp 8, results='asis', fig.height=4, fig.width=9} -->
<!-- #all on one plot -->
<!-- p_8 <- ggplot(data=arr_all)+ -->
<!--     geom_smooth( aes(x =time, y = mean_both, colour=strata), linewidth=0.5, linetype="dashed")+ -->
<!--     geom_smooth( data=arr_comb_1,aes(x =time, y = mean_both, colour=strata), linewidth=1)+ -->
<!--     geom_ribbon(data=arr_comb_1, aes(ymin = l_ci_both, ymax = u_ci_both, x=time),fill= "#377EB8",alpha=0.1)+  -->
<!--     scale_x_continuous("Time (years)", breaks = c(0,365.25,365.25*2, 365.25*3, 365.25*4, 365.25*5, 365.25*6), -->
<!--                  labels = c(0,1,2,3,4,5,6), limits=c(0,2300)) + -->
<!--     scale_y_continuous ("Absolute risk reduction (%)") + -->
<!--     scale_color_discrete(palette=c("#377EB8","#E41A1C",  "#4DAF4A" ,"#984EA3", "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))+ -->
<!--     theme_bw()+ -->
<!--     ggtitle(paste("Absolute risk reduction of recurrent stroke, HR for competing risk of death = 0.8"))+ -->
<!--     labs(color="Stratum") -->
<!-- p_8 -->
<!-- ggsave("ARR plots with varying competing risks/Competing risk ratio 0.8, combined.png") -->
<!-- ``` -->

<!-- #### Compare plots with different competing risks -->

<!-- ```{r combine plots with different CR, fig.height=6, fig.width=10} -->
<!-- p_real+p_1+p_9+p_8+plot_layout(guides="collect", axes="collect")& theme(title=element_text(size=6), axis.title=element_text(size=8)) -->
<!-- ``` -->